---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zimsSRGa

<!-- badges: start -->
<!-- badges: end -->


zimsSRGa can be used to run the main demographic analyses: survival, reproduction and growth analytics from zims data.
The data need to be cleaned before using this package.



## Installation

First you will need install paramDemo on your computer from [GitHub](https://github.com/) with:

```{r, echo = TRUE, eval = FALSE}
# install.packages("devtools")
# library(devtools)
install_git("https://github.com/fercol/paramDemo", subdir = "pkg/")


```


You can install the development version of zimsSRGa from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("species360/zimsSRGa")
```

## Documentation

Full documentation website on: XXXXXXXXXXXXXX

You can also open the documentation locally on your machine using
``` r
path <- system.file("site", "index.html", package = "zimsSRGa")
browseURL(path)
```

## Tutorial for survival


```{r, echo = TRUE, eval = FALSE}
ExtractDate ="2024-08-29"
Taxa = "Mammalia"
List_species = list(Mammalia = c("Capra sibirica", "Cuon alpinus", "Nanger granti"))

#Filters -----------------------------------------------------------------------
# Earliest date to include records
MinDate <- "1980-01-01"
# Earliest birth date to include records
MinBirthDate <- "1900-01-01"
# Whether to include only Global individuals
Global = TRUE
# Birth Type of Animals: "Captive", "Wild" or "All"
BirthType = "Captive"
# Minimum number of individuals to run the taxon profile
MinN <- 50
# Maximum threshold in the longevity distribution to use
MaxOutl <- 99 
# Minimum number of Institutions that hold individuals from one species
MinInstitution = 2 
# Maximum uncertainty accepted for birth dates, in days
UncertBirth = 365
# Maximum uncertainty accepted for death dates, in days
UncertDeath = 365
# Maximum possible age
MaxAge = 120
# Age when to start the analysis
MinAge = 0


# Survival Models ------------------------------------------------------------
# Survival Models to run: "GO", "LO", "EX" or/and "WE"
ModelsSur <- c("GO", "LO")
# Shape of the survival model: "simple", "bathtub" or "Makeham"
Shape = "bathtub"

# Number of CPUS:
ncpus <- 4
# MCMC settings:
niter <- 10000
burnin <- 3001
thinning <- 20
nchain <- 3

# Conditions to run the survival analysis
MinNSur = 50 #Minimum number of individuals
MaxNSur = 1000 #Maximum number of individuals
MinLx = 0.1  #Minimum survivorship reach by life table (Kaplan-Meier estimator)
MinBirthKnown = 0.3 #Minimum proportions of known birth dates (within a month)
#Goodness of fit
MinMLE = 0.1 #Minimum survivorship at Mean life expectancy
MaxLE = 2     #Maximum remaining life expectancy at max age

# Table with species that require rerruning with higher outlier level:
MaxOutlev <- read.csv(glue("{AnalysisDir}MaxOutlierLevel.csv"), header = TRUE,
                      stringsAsFactors = FALSE)
spOutLev <- MaxOutlev$Specie


# Run survival Analysis---------------------------------------------------------
## Loop over species
for (species in List_species[[Taxa]]){
  print(species)
  if (species %in% spOutLev) {MaxOutl1 <- 99.9
  }else{MaxOutl1 <- MaxOutl}
  Dataspe <- select_species(Species, Animal, Data[[Taxa]]$Collection, UncertBirth = UncertBirth,
                            BirthType = BirthType,
                            MinDate = MinDate , ExtractDate = ExtractDate,
                            Global = Global) 
  if(nrow(Dataspe$data)>0){
    out = list()
    ## Loop over sexes
    for (sx in c("Male", "Female")){
      print(sx)
      #Check for gaps in longevity distribution to define a threshold 
      #after which individuals will be deleted (possible outiers or errors in data)
      sexDat <- select_Longthreshold( Dataspe$data,  SexCats = sx, 
                                      PlotDir= glue::glue("{PlotDir}/"),
                                      MinN = MinN ,
                                      PlotName = glue::glue("{Taxa}_{species}_{sx}") )
      OutlLev1 = min(sexDat$summar$GapThresh,MaxOutl, na.rm = T)
      
      
      if(nrow(sexDat$data)>0){ 
        out[[sx]] <-  Sur_main(DataCore = sexDat$data,  DeathInformation =  Data[[Taxa]]$DeathInformation,
                               BirthType = BirthType, LastDead = T,
                               PlotDir = glue::glue("{PlotDir}/"),
                               MaxAge = MaxAge,
                               Models = ModelsSur, Shape= Shape, 
                               OutlLev1 =OutlLev1,
                               MinMLE = MinMLE, MaxLE =  MaxLE,
                               MinDate = MinDate, MinNSur = MinNSur, MaxNSur = MaxNSur, 
                               MinInstitution = MinInstitution,UncertDeath= UncertDeath,
                               MinLx = MinLx , MinBirthKnown = MinBirthKnown, 
                               niter = niter, burnin = burnin, thinning = thinning, nchain = nchain, 
                               ncpus = ncpus, PlotName = glue("{Taxa}_{species}_{sx}") )
        print(out$summary$error)
      }
    }
  }
}

#Plot predictions and fit from Basta survival model
plot(out$bastaRes, plot.type = 'gof')
```



