---
title: "flat_additional.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
library(assertthat)
library(tidyverse)
library(glue)

options(rmarkdown.html_vignette.check_title = FALSE)

```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# Main survival Analysis

`Sur_main()` runs the survival analysis (see next function : `Sur_ana()`) and estimates associated key survival metrics. In addition to the arguments of `Sur_ana()`, `Sur_main()` includes the main following arguments:

* `DataCore` the core data of the species from ZIMS including at least the following columns *AnimalAnonID*, *BirthDate*, *DepartDate*, *EntryDate*, *MaxBirthDate*, *MinBirthDate*, *EntryType*, *DepartType*, *FirstHoldingInstitution*, *LastHoldingInstitution*, *SexType*, and *BirthType*.
* `DeathInformation` data.frame including at least the following columns *AnimalAnonID* and *RelevantDeathInformationType*
* `CalculateMetricsFrom` indicates which metrics should be calculated from which data: "Raw", "Kaplan-Meier", "Model".
* `BirthType_sel`: The birth type to be selected: Captive, Wild, or All
* `PlotDir`: Directory to save the plots of the selected model

It returns a list including, depending on `CalculateMetricsFrom`:
* a summary of the data selected
* The Kaplan-Meier estimator
* the BaSTA fit of the selected model
* the DIC table comparing the fit  of the different models
* the remaining life expectancy per age
* key survival metrics
* Statistics to check the fit of the selected model

```{r function-Sur_main}
#' Main survival analysis
#' 
#' Run the survival analysis and calculates key survival metrics
#'
#' @param DataCore \code{data.frame} including at least the following columns *AnimalAnonID*, *BirthDate*, *DepartDate*, *EntryDate*, *MaxBirthDate*, *MinBirthDate*, *EntryType*, *DepartType*, *FirstHoldingInstitution*, *LastHoldingInstitution*, *SexType*, and *BirthType*.
#' @param DeathInformation  \code{data.frame} including at least the following columns *AnimalAnonID* and *RelevantDeathInformationType*.
#' @param BirthType_sel \code{character} The birth type to be selected: Captive, Wild, or All.
#' @param CalculateMetricsFrom \code{vector of characters} indicates which metrics should be calculated from which data: "Raw", "Kaplan-Meier", "Model".
#' @param Models \code{vector of characters} names of the BaSTA Models to run: "GO", "EX", "LO" and/or "WE". see ?basta for more information. 
#' @param Shape \code{character} Shape of the BaSTA model: "simple", "Makeham", "bathtub".  see ?basta for more information.
#' @param MinAge \code{numeric} Ages at which the survival analysis should start in years.  see ?basta for more information.
#' @param MaxAge \code{numeric} Maximum possible age in years. Only used for model predictions. This argument is not used to select data.
#' @param OutlLev1 \code{numeric} Start threshold to select individuals based on the longevity distribution: 100%, 99.9, 99 or 95%. This number must decrease when many errors are expected in longevity data.
#' @param OutlLevmin \code{numeric} Minimum threshold to select individuals based on the longevity distribution: 100%, 99.9, 99 or 95%. 
#' @param UncertDeath \code{numeric}: Maximum uncertainty accepted for death date, in days.
#' @param MinDate \code{character 'YYYY-MM-DD'} Earlier date to include data.
#' @param MinNSur \code{numeric} Minimum number of individual records needed to run the survival analysis.
#' @param MaxNSur \code{numeric} Maximum number of individual records allowed to run the survival analysis.
#' @param MinInstitution \code{numeric} Minimum number of institutions that should hold records to run the survival analysis.
#' @param MinBirthKnown  \code{numeric} between 0 and 1. Minimum proportion of individuals with a known birth month in the data to run the survival analysis.
#' @param MinLx  \code{numeric} Value used for longevity threshold and for checks. between 0 and 1. Minimum reached survivorship from the raw Kaplan-Meier analysis. This number avoids running survival analysis if there are too few dead individuals in the data. Lower is better.
#' @param MinMLE \code{numeric} Value used for checks. Minimum survivorship allowed at mean life expectancy. Between 0 and 1.
#' @param MaxLE \code{numeric} Value used for checks. Maximum remaining life expectancy at last observed age. In years.
#' @param niter  \code{numeric}. Parameter to run the model: number of MCMC iterations. see ?basta for more information.
#' @param burnin  \code{numeric} Parameter to run the model: number of iterations removed so that the model has time to converge. see ?basta for more information.
#' @param thinning  \code{numeric} Parameter to run the model: number of iterations to run before saving a set of parameters. see ?basta for more information.
#' @param nchain  \code{numeric} Parameter to run the model: Number of chains to run.
#' @param ncpus  \code{numeric} Number of computer cores to use. 
#' @param PlotDir \code{character} Directory to save the plots. If NULL, no plot is saved
#' @param PlotName \code{character} Name used to save the plots.
#' @param LastDead  \code{logical} Whether the longest lived individuals should be considered dead.
#' 
#' @return A named list with one element per value of `MinAge`, named
#'   `from<MinAge>` (e.g. `from0`). Each element is itself a list containing
#'   `summary`, `metrics`, `bastaRes`, `DICmods`, `KM_estimator`, `bastatab`,
#'   `relex_from0`, `Sur1`, `Sur5`, `Sur1m`, and `check`. Some components may be
#'   `NULL` depending on `CalculateMetricsFrom` and whether model fitting was successful
#' * a summary of the data used:
#'    * NSelect: Number of individuals selected from filters
#'    * NUncertdeath: Number of individuals selected after filter uncertainty in death
#'    * NBasta: Number of data (individuals) selected for the BaSTA/survival analysis
#'    * Ndead:Number of selected individuals with known age of death used in the BaSTA/survival analysis
#'    * Nrc: Number of selected right-censored individuals
#'    * N8090: Number of selected individuals born between 1980 and 1990
#'    * N9000: Number of selected individuals born between 1990 and 2000
#'    * N0010: Number of selected individuals born between 2000 and 2010
#'    * N1020: Number of selected individuals born between 2010 and 2020
#'    * N2030: Number of selected individuals born between 2020 and 2030
#'    * QBD10: 10% Quantile of birth date distribution among selected individuals
#'    * QBD50: Median of birth date distribution among selected individuals
#'    * QBD90: 90% Quantile of birth date distribution among selected individuals
#'    * BDincert: average uncertainty in birth date: in days among selected individuals
#'    * maxAge: Maximum age of selected known age individuals
#'    * maxAlive:Maximum number of years spent ex situ among selected individuals
#'    * lxMin: Minimum survivorship reached with the raw Kaplan-Meier model
#'    * outLev: threshold selected for the distribution of longevity: 100%, 99.9%, 99% or 95%
#'    * analyzed: logical indicating whether the BaSTA survival model was fitted.
#'    * Nerr, error: error code and message when the model was not fitted or when a goodness-of-fit check failed.
#' * the basta fit of the best model
#' * the DIC table comparing the different fit of the Models
#' * Key survival metrics including Mean life expectancy (MLE & Ex), median life expectancy (L50) and  age at which 90% of the individual died (Longevity = L90) estimated from raw data, the Kaplan-Meier estimator, and the survival model. Estimates include ages from birth or from age at sexual maturity (if given). First year survival, First month survival, Entropy (H and Epx = -log(H)), coefficient of variation (CV) and Gini coefficient (G) are also estimated from the Kaplan-Meier estimator and the survival model.
#' * Checks of the fit for the selected survival model:
#'     * Gof_KM_coeff1: Percentage of age points where the KM estimator is outside of the 95% CI of the survivorship estimated from the survival model.
#'     * Gof_KM_coeff2: Maximum sum of same sign residuals between KM estimator and survival model
#'     * Gof_Martingale: Goodness of fit using Martingale residuals
#'     * LxatMLE: Estimated survivorship at mean life expectancy
#'     * LEmaxOage: Life expectancy at max observed age
#'     * KMMinLx: Minimum survivorship reached by the Kaplan-Meier estimator
#' * The remaining life expectancy per age (relex_from0)
#' * The probability to live 5 years more (Sur5)
#' * Age-specific yearly survival  (Sur1)
#' * Monthly survival  (Sur1m)
#' * If PlotDir is filled, 2 plots are produced: 
#'     * One showing the convergence of chains, together with estimated survivorship and survival and mortality rates.
#'     * The second showing remaining life-expectancy, and the age-specific probabilities to live 5 years and 1 year more.
#'
#' @export
#' @import assertthat tidyverse dplyr
#' 
#' @examples
Sur_main <- function(DataCore,   DeathInformation, 
                     CalculateMetricsFrom = c("Raw", "Kaplan-Meier", "Model"),
                     BirthType_sel = "All",
                     Models = "GO", Shape= "simple",
                     MinAge =0, MaxAge=120, OutlLev1 = NA,OutlLevmin = 95,
                     UncertDeath=365,
                     MinDate = "1980-01-01", MinNSur = 50, MaxNSur = NULL,
                     MinInstitution = 2,
                     MinLx = 0.1, MinBirthKnown = 0.3,
                     MinMLE = 0.1, MaxLE = 2, 
                     LastDead = FALSE,
                     niter = 25000, burnin = 5001, thinning = 20, nchain = 3, 
                     ncpus = 2, PlotDir = NULL, PlotName = '') {
  
  # Check correct format for inputs -----------------------------------------------------------------------
  assert_that(is.data.frame(DataCore))
  assert_that(DataCore  %has_name% c("AnimalAnonID", "BirthDate", "DepartDate",
                                     "EntryDate", "MaxBirthDate", "MinBirthDate",
                                     "EntryType", "DepartType", "FirstHoldingInstitution", 
                                     "LastHoldingInstitution", "SexType", "BirthType"))
  assert_that(is.data.frame(DeathInformation))
  assert_that(DeathInformation %has_name% c("AnimalAnonID","RelevantDeathInformationType"))
 assert_that(is.character(BirthType_sel), length(BirthType_sel) == 1)
  assert_that(BirthType_sel %in% c("Captive", "Wild", "All"))

  assert_that(is.character(Models))
  assert_that(all(Models %in% c("", "GO", "EX", "LO", "WE")))

  assert_that(is.character(CalculateMetricsFrom))
  assert_that(all(CalculateMetricsFrom %in% c("Raw", "Kaplan-Meier", "Model")))

  assert_that(is.character(Shape), length(Shape) == 1)
  assert_that(Shape %in% c("", "simple", "bathtub", "Makeham"))

  assert_that(is.numeric(MinAge), length(MinAge) >= 1)
  assert_that(is.numeric(MaxAge), length(MaxAge) == 1)
  assert_that(all(MinAge < MaxAge))

  if (!is.na(OutlLev1)) {
    assert_that(is.numeric(OutlLev1), length(OutlLev1) == 1)
    assert_that(OutlLev1 <= 100)
  } else {
    OutlLev1 <- 100
  }

  assert_that(is.numeric(OutlLevmin), length(OutlLevmin) == 1)
  assert_that(OutlLevmin <= 100)
  assert_that(OutlLevmin >= 0)

  assert_that(is.numeric(UncertDeath), length(UncertDeath) == 1)
  assert_that(UncertDeath >= 0)

  MinDate <- lubridate::as_date(MinDate)
  assert_that(!is.na(MinDate), msg = "MinDate must be a valid date in format YYYY-MM-DD")

  assert_that(is.numeric(MinNSur), length(MinNSur) == 1)
  if (!is.null(MaxNSur)) assert_that(is.numeric(MaxNSur), length(MaxNSur) == 1)

  assert_that(is.numeric(MinInstitution), length(MinInstitution) == 1, MinInstitution > 0)

  assert_that(is.numeric(MinLx), length(MinLx) == 1)
  assert_that(MinLx >= 0 && MinLx <= 1)

  assert_that(is.numeric(MinMLE), length(MinMLE) == 1)
  assert_that(MinMLE >= 0 && MinMLE <= 1)

  assert_that(is.numeric(MaxLE), length(MaxLE) == 1)

  assert_that(is.numeric(MinBirthKnown), length(MinBirthKnown) == 1)
  assert_that(MinBirthKnown >= 0 && MinBirthKnown <= 1,
              msg = "MinBirthKnown must be a proportion between 0 and 1")

  assert_that(is.logical(LastDead), length(LastDead) == 1)

  assert_that(is.numeric(niter), length(niter) == 1, niter > 0)
  assert_that(is.numeric(burnin), length(burnin) == 1, burnin > 0, burnin < niter)
  assert_that(is.numeric(thinning), length(thinning) == 1, thinning > 0, thinning < niter)
  assert_that(is.numeric(nchain), length(nchain) == 1, nchain > 0)
  assert_that(is.numeric(ncpus), length(ncpus) == 1, ncpus > 0)

  assert_that(is.character(PlotName), length(PlotName) == 1)

  if (!is.null(PlotDir)) {
    checkmate::assert_directory_exists(PlotDir)
  }

  
  # Initialize outputs ---------------------------------------------------------------------------------
  out = list()
  
  # Select individuals with correct birth type----------------------------------------------------------
  if (BirthType_sel != "All"){
    Data<- DataCore %>%
      dplyr::filter(stringr::str_detect(BirthType, pattern = BirthType_sel))
  }else{Data<- DataCore}
  
  # Run Survival Analysis-------------------------------------------------------------------------------
  for(j in seq_along(MinAge)){
     assert_that(MinAge[j] < MaxAge)

    out[[paste0("from",MinAge[j])]]<- Sur_ana(Data, DeathInformation = DeathInformation,
                                              OutlLev1 = OutlLev1, OutlLevmin = OutlLevmin, 
                                              CalculateMetricsFrom = CalculateMetricsFrom,
                                              Models = Models, Shape = Shape,
                                              MinDate = MinDate, UncertDeath = UncertDeath, 
                                              LastDead = LastDead, MinAge = MinAge[j],
                                              MinNSur = MinNSur, MaxNSur = MaxNSur,
                                              MinLx = MinLx,
                                              MinInstitution = MinInstitution,
                                              MinBirthKnown = MinBirthKnown, 
                                              niter = niter, burnin = burnin, thinning = thinning, 
                                              nchain = nchain, ncpus = ncpus) 
    out[[paste0("from",MinAge[j])]] <- append( out[[paste0("from",MinAge[j])]], 
                                               list(relex_from0 = NULL, Sur1 = NULL, Sur5 = NULL, Sur1m = NULL, check = NULL))
    
    #Estimate key survival metrics
    if (out[[paste0("from",MinAge[j])]]$summary$analyzed) {
      out[[paste0("from",MinAge[j])]]  <-Sur_out(out[[paste0("from",MinAge[j])]], 
                                                 MaxAge=MaxAge,ncpus = ncpus, MinAge = MinAge[j],
                                                 PlotDir = PlotDir, PlotName = paste0(PlotName,"_",MinAge[j]),
                                                 MinLx = MinLx, MinMLE = MinMLE, MaxLE = MaxLE
      )
    }
  }
  
  return(out)
}
```

```{r examples-Sur_main}
data(toy_core)
data(toy_deathinformation)
out <- Sur_main(toy_core, DeathInformation = toy_deathinformation, BirthType_sel = "All",
                Models = "GO", Shape = "simple",
                niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
```

```{r tests-Sur_main}
test_that("Sur_main works", {
  data(toy_core)
  data(toy_deathinformation)
  out <- Sur_main(toy_core, DeathInformation = toy_deathinformation, BirthType_sel = "Captive", 
                  Models = "LO", Shape = "bathtub",
                  niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
  expect_named(out, c("from0"))
  expect_named(out$from0, c('summary', 'metrics', 'bastaRes', 'DICmods', 'KM_estimator', 'bastatab', 'relex_from0',  'Sur1', 'Sur5', 'Sur1m', 'check'))
  expect_named(out$from0$summary, c('NSelect', 'NUncertdeath', 'NBasta', 'Ndead', 'Nrc', 'N8090', 'N9000', 'N0010', 'N1020', 'N2030', 'QBD10', 'QBD50', 'QBD90', 'BDincert', 'maxAge', 'maxAlive', 'lxMin', 'outLev', 'model', 'analyzed', 'Nerr', 'error'))
  expect_named(out$from0$DICmods, c("Models", "DIC"))
  out <- Sur_main(toy_core, DeathInformation = toy_deathinformation, BirthType_sel = "Captive", 
                  Models = "LO", Shape = "bathtub", MaxNSur = 10, MinNSur = 1,
                  niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
  
})
```

# Survival Analysis

This function runs BaSTA survival models on the data after checking conditions for correct convergence of the models. It includes the following arguments:
* `Data`including at least the following columns *AnimalAnonID*, *BirthDate*, *DepartDate*, *EntryDate*, *MaxBirthDate*, *MinBirthDate*, *EntryType*, *DepartType*, *FirstHoldingInstitution*, *LastHoldingInstitution*, *SexType*, and *BirthType*.
* `DeathInformation` data.frame including at least the following columns *AnimalAnonID* and *RelevantDeathInformationType*.
* `CalculateMetricsFrom` indicates which metrics should be calculated from which data: "Raw", "Kaplan-Meier", "Model".
* `Models` names of the BaSTA Models to run: "GO", "EX", "LO" and/or "WE". see ?basta for more information. Default = "GO".
* `Shape` shape of the BaSTA model: "simple", "Makeham", "bathtub".  see ?basta for more information. Default = "simple".

The user should define different filters to select data and run the survival analyses. The main ones are: 
* `UncertDeath` Maximum uncertainty accepted for death date, in days.
* `MinNSur` Minimum number of individual records needed to run the survival analysis.
* `MaxNSur` Maximum number of individual records to run the survival analysis.
* `MinLx` Minimum reached survivorship from the raw Kaplan-Meier analysis needed to run the survival analysis.
* `MinBirthKnown` Minimum proportion of individuals with a known birth month in the data.
* `MinInstitution` Minimum number of institutions that should hold records to run the survival analysis.

It returns a list including and depending on `CalculateMetricsFrom`:
* a summary of the data used
* The Kaplan-Meier estimator and associated key metrics
* Key survival metrics from raw data
* the BaSTA fit of the best model
* the DIC table comparing the different fit of the Models


```{r function-Sur_ana}
#' Survival Analysis
#' 
#' Run BaSTA Models on the data after checking conditions for correct convergence of the Models
#'
#' @param Data \code{data.frame} including at least the following columns *AnimalAnonID*, *BirthDate*, *DepartDate*, *EntryDate*, *MaxBirthDate*, *MinBirthDate*, *EntryType*, *DepartType*, *FirstHoldingInstitution*, *LastHoldingInstitution*, *SexType*, and *BirthType*.
#' @param DeathInformation  \code{data.frame} including at least the following columns *AnimalAnonID* and *RelevantDeathInformationType*
#' @param CalculateMetricsFrom \code{vector of characters} indicates which metrics should be calculated from which data: "Raw", "Kaplan-Meier", "Model".
#' @param Models \code{vector of characters} names of the BaSTA models to run: "GO", "EX", "LO" and/or "WE". see ?basta for more information.
#' @param Shape \code{character} Shape of the BaSTA model: "simple", "Makeham", "bathtub".  see ?basta for more information.
#' @param MinAge \code{numeric} Ages at which the survival analysis should start, in years.  see ?basta for more information.
#' @param OutlLev1 \code{numeric} Start threshold to select individuals based on the longevity distribution: 100%, 99.9, 99 or 95%. This number must decrease when many errors are expected in longevity data.
#' @param OutlLevmin \code{numeric} Minimum threshold to select individuals based on the longevity distribution: 100%, 99.9, 99 or 95%. 
#' @param MinDate \code{character 'YYYY-MM-DD'} Earlier date to include data.
#' @param UncertDeath \code{numeric}: Maximum uncertainty accepted for death date, in days.
#' @param MinNSur \code{numeric} Minimum number of individual records needed to run the survival analysis.
#' @param MaxNSur \code{numeric} Maximum number of individual records to run the survival analysis.
#' @param MinLx  \code{numeric} Value used for longevity threshold and for checks. between 0 and 1. Minimum reached survivorship from the raw Kaplan-Meier analysis. This number avoids running survival analysis if there are too few dead individuals in the data. Lower is better.
#' @param MinInstitution \code{numeric} Minimum number of institutions that should hold records to run the survival analysis. 
#' @param MinBirthKnown  \code{numeric} between 0 and 1. Minimum proportion of individuals with a known birth month in the data to run the survival analysis.
#' @param niter  \code{numeric}. Parameter to run the model: number of MCMC iterations. see ?basta for more information.
#' @param burnin  \code{numeric} Parameter to run the model: number of iterations removed so that the model has time to converge. see ?basta for more information.
#' @param thinning  \code{numeric} Parameter to run the model: number of iterations to run before saving a set of parameters. see ?basta for more information.
#' @param nchain  \code{numeric} Parameter to run the model: Number of chains to run.
#' @param ncpus  \code{numeric} Number of computer cores to use. 
#' @param LastDead  \code{logical} Whether the longest lived individuals should be considered dead.
#'
#' @return A list with components `summary`, `metrics`, `bastaRes`, `DICmods`,
#'   `KM_estimator`, and `bastatab`. Some components may be `NULL` depending on
#'   `CalculateMetricsFrom` and whether model fitting succeeded.
#' * a summary of the data used:
#'    * NSelect: Number of individuals selected from filters
#'    * NUncertdeath: Number of individuals selected after filter uncertainty in death
#'    * NBasta: Number of data (individuals) selected for the BaSTA/survival analysis
#'    * Ndead:Number of selected individuals with known age of death used in the BaSTA/survival analysis
#'    * Nrc: Number of selected right-censored individuals
#'    * N8090: Number of selected individuals born between 1980 and 1990
#'    * N9000: Number of selected individuals born between 1990 and 2000
#'    * N0010: Number of selected individuals born between 2000 and 2010
#'    * N1020: Number of selected individuals born between 2010 and 2020
#'    * N2030: Number of selected individuals born between 2020 and 2030
#'    * QBD10: 10% Quantile of birth date distribution among selected individuals
#'    * QBD50: Median of birth date distribution among selected individuals
#'    * QBD90: 90% Quantile of birth date distribution among selected individuals
#'    * BDincert: average uncertainty in birth date: in days among selected individuals
#'    * maxAge: Maximum age of selected known age individuals
#'    * maxAlive:Maximum number of years spent ex situ among selected individuals
#'    * lxMin: Minimum survivorship reached with the raw Kaplan-Meier model
#'    * outLev: threshold selected for the distribution of longevity: 100%, 99.9%, 99% or 95%
#'    * analyzed: a logical indicated if the basta survival model was performed
#'    * If the basta survival model was not performed, an error and its number (Nerr) are returned: The possibility for  this function are: 2/Nuncertdeath < MinNSur; 3/ lxMin> 0.75; 4/NBasta = 0; 5/ %known births < MinBirthKnown; 6/Number of institutions < MinInstitution; 7/Nbasta < MinNSur; 8/Nbasta > MaxNSur; 9/no DIC from Basta; 10/Gof martingale does not fit
#' * the Kaplan-Meier table
#' * Key survival metrics including Mean life expectancy (MLE & Ex), median life expectancy (L50) and  age at which 90% of the individual died (Longevity = L90) estimated from raw data and from the Kaplan-Meier Estimator. Estimates include ages from birth or from age at sexual maturity (if given). First year survival, First month survival, Entropy (H and Epx = -log(H)), coefficient of variation (CV) and Gini coefficient (G) are also estimated from the Kaplan-Meier estimator.
#'* the basta fit of the best model
#'* the DIC table comparing the different fit of the Models
#' 
#' @export
#' @importFrom lubridate year
#' @importFrom BaSTA basta
#' @importFrom stats median quantile sd
#'
#' @examples
Sur_ana <- function(Data, DeathInformation, 
                    CalculateMetricsFrom = c("Raw", "Kaplan-Meier", "Model"), 
                    Models = "GO", Shape = "simple", 
                    MinAge = 0, OutlLev1 = 100,OutlLevmin = 95,
                    MinDate = "1980-01-01", MinNSur = 50, MaxNSur = NULL,
                    UncertDeath=365, MinLx=0.1,
                    MinBirthKnown = 0.3, MinInstitution = 1,
                    LastDead = FALSE,
                    niter = 25000, burnin = 5001, thinning = 20, nchain = 3,
                    ncpus = 2) {
  
  # Check correct format for inputs ---------------------------------------------
  required_cols <- c(
    "AnimalAnonID", "BirthDate", "DepartDate", "EntryDate",
    "MaxBirthDate", "MinBirthDate", "EntryType", "DepartType",
    "FirstHoldingInstitution", "LastHoldingInstitution",
    "SexType", "BirthType", "DeathUncertainty", "GlobalStatus"
  )
  
  assert_that(is.data.frame(Data))
  assert_that(Data %has_name% required_cols)
  
  assert_that(is.data.frame(DeathInformation))
  assert_that(DeathInformation %has_name% c("AnimalAnonID", "RelevantDeathInformationType"))
  
  assert_that(is.character(CalculateMetricsFrom))
  assert_that(length(CalculateMetricsFrom) >= 1)
  assert_that(all(CalculateMetricsFrom %in% c("Raw", "Kaplan-Meier", "Model")))
  
  assert_that(is.character(Models))
  assert_that(length(Models) >= 1)
  assert_that(all(Models %in% c("", "GO", "EX", "LO", "WE")))
  
  assert_that(is.character(Shape))
  assert_that(length(Shape) == 1)
  assert_that(Shape %in% c("", "simple", "bathtub", "Makeham"))
  
  assert_that(is.numeric(MinAge))
  assert_that(length(MinAge) == 1)
  assert_that(MinAge >= 0)
  
  assert_that(is.numeric(OutlLev1))
  assert_that(length(OutlLev1) == 1)
  assert_that(OutlLev1 >= 0 && OutlLev1 <= 100)
  
  assert_that(is.numeric(OutlLevmin))
  assert_that(length(OutlLevmin) == 1)
  assert_that(OutlLevmin >= 0 && OutlLevmin <= 100)
  assert_that(OutlLev1 >= OutlLevmin)
  
  assert_that(is.numeric(UncertDeath))
  assert_that(length(UncertDeath) == 1)
  assert_that(UncertDeath >= 0)
  
  MinDate <- lubridate::as_date(MinDate)
  assert_that(!is.na(MinDate), msg = "MinDate must be a valid date in format 'YYYY-MM-DD'")
  
  assert_that(is.numeric(MinNSur))
  assert_that(length(MinNSur) == 1)
  assert_that(MinNSur > 0)
  
  if (!is.null(MaxNSur)) {
    assert_that(is.numeric(MaxNSur))
    assert_that(length(MaxNSur) == 1)
    assert_that(MaxNSur > 0)
    assert_that(MaxNSur >= MinNSur)
  } else {
    MaxNSur <- nrow(Data)
  }
  
  assert_that(is.numeric(MinLx))
  assert_that(length(MinLx) == 1)
  assert_that(MinLx >= 0 && MinLx <= 1)
  
  assert_that(is.numeric(MinBirthKnown))
  assert_that(length(MinBirthKnown) == 1)
  assert_that(MinBirthKnown >= 0 && MinBirthKnown <= 1,
              msg = "MinBirthKnown must be a proportion between 0 and 1")
  
  assert_that(is.numeric(MinInstitution))
  assert_that(length(MinInstitution) == 1)
  assert_that(MinInstitution > 0)
  
  assert_that(is.logical(LastDead))
  assert_that(length(LastDead) == 1)
  
  assert_that(is.numeric(niter))
  assert_that(length(niter) == 1)
  assert_that(niter > 0)
  
  assert_that(is.numeric(burnin))
  assert_that(length(burnin) == 1)
  assert_that(burnin > 0)
  assert_that(burnin < niter)
  
  assert_that(is.numeric(thinning))
  assert_that(length(thinning) == 1)
  assert_that(thinning > 0)
  assert_that(thinning < niter)
  
  assert_that(is.numeric(nchain))
  assert_that(length(nchain) == 1)
  assert_that(nchain > 0)
  
  assert_that(is.numeric(ncpus))
  assert_that(length(ncpus) == 1)
  assert_that(ncpus > 0)
  
  # Initialize outpus ----------------------------------------------------------
  summar = list(
    NSelect =  nrow(Data), NUncertdeath = 0,  
    NBasta = 0, Ndead = 0, Nrc = 0,
    N8090 =0, N9000 = 0, N0010 = 0, N1020 = 0, N2030 = 0,
    QBD10 = 0, QBD50 = 0, QBD90 = 0, BDincert = 0,
    maxAge = NULL, maxAlive = NULL, 
    lxMin = NULL, outLev = NULL, model ="",
    analyzed = FALSE, Nerr = 0,  error ="")
  metrics =  tibble::tibble(Data = character(0),
                    firstage = numeric(0) ,
                    param = character(0),
                    stat =character(0),
                    value = numeric(0))
  bastaRes = NULL
  bastatab = NULL
  DICmods = NULL
  KMest = NULL
  # Remove individuals with uncertainty in death date----------------------------
  Data <- Data%>%
    dplyr::filter((DeathUncertainty < UncertDeath)%>% tidyr::replace_na(TRUE))
  summar$NUncertdeath =nrow(Data)
  
  # Check for possible gaps in longevity distribution and deduce maximum longevity threshold specifically on the survival dataset
  if(nrow(Data)> MinNSur){
    Dat <- select_Longthreshold(Data, MinN = MinNSur)
    # deduce maximum longevity threshold
    outlLev = max(Dat$summar$GapThresh,OutlLevmin)
    
    # Find the maximum longevity threshold (min = 95%) for which lxmin > MinLx----
    summar$lxMin <- 1
    outLev2 = min(OutlLev1, outlLev)
   Datar <- Data%>%
        dplyr::mutate(
          deparAge = (DepartDate - BirthDate) / 365.25,
          entryAge = (EntryDate - BirthDate) / 365.25)%>%
     dplyr::filter(deparAge>=MinAge)
   
    while(summar$lxMin > MinLx & outLev2 >= OutlLevmin){
      summar$outLev = outLev2
      if (summar$outLev ==100){
        data_sel <-  Datar
      }else{
        data_sel <-  Datar%>%
          dplyr::filter(!!sym(paste0("above", summar$outLev))==0)
      }
      data_sel <- data_sel%>%
        dplyr::filter(deparAge>=MinAge)
      
      if (!all(data_sel$DepartType == "C")) {
        #Calculate Kaplan-Meier table
        rawPLE <- Sur_ple(data_sel)
        summar$lxMin <- rawPLE$ple[nrow(rawPLE)-1]
      }
      if(summar$outLev == 100){outLev2 = 99.9}
      if(summar$outLev == 99.9){outLev2 = 99}
      if(summar$outLev == 99){outLev2 = 95}
      if(summar$outLev == 95){outLev2 = 90}
    }

    # Estimate key survival metrics from Raw data and from Kaplan-Meier table----
    if(summar$lxMin < 0.75){
      #from Age at sexual maturity
       #from birth
      if ("Raw" %in% CalculateMetricsFrom){
        metrics = rbind(metrics,
                        tibble::tibble(Data = 'Raw',
                               firstage = MinAge,
                               param = c("S1m","S1y", "L50", "MLE", "L90", "L99", "Max_longevity"),
                               stat = c("value"),
                               value = c(1-(length( Datar$deparAge[Datar$DepartType == "D"& Datar$deparAge<= 1/12])/
                                           length( Datar$deparAge[Datar$DepartType == "D"]))%>%as.numeric(),
                                         1-(length( Datar$deparAge[Datar$DepartType == "D"& Datar$deparAge<= 1])/
                                           length( Datar$deparAge[Datar$DepartType == "D"]))%>%as.numeric(),
                                         median( Datar$deparAge[Datar$DepartType == "D"])%>%as.numeric(),
                                         mean( Datar$deparAge[Datar$DepartType == "D"])%>%as.numeric(),
                                         quantile( Datar$deparAge[Datar$DepartType == "D"],0.9)%>%as.numeric(),
                                         quantile( Datar$deparAge[Datar$DepartType == "D"],0.99)%>%as.numeric(),
                                         max( Datar$deparAge[Datar$DepartType == "D"])%>%as.numeric()
                               )))
        if(length(Datar$BirthDate[Datar$GlobalStatus == "Alive"])>0){
          maxcoh = min(year(Datar$BirthDate[Datar$GlobalStatus == "Alive"]))
          Datamax = Datar%>%dplyr::filter(year(BirthDate)<maxcoh, DepartType == "D")
          if(nrow(Datamax)>0){
            metrics = rbind(metrics,
                            tibble::tibble(Data = 'Raw_Deadcoh',
                                   firstage = MinAge,
                                   param = c("S1m","S1y","L50", "MLE", "L90", "L99", "Max_longevity", "N"),
                                   stat = c("value"),
                                   value = c(1-(length(Datamax$deparAge[Datamax$DepartType == "D"& Datamax$deparAge<= 1/12])/
                                           length( Datamax$deparAge[Datamax$DepartType == "D"]))%>%as.numeric(),
                                         1-(length( Datamax$deparAge[Datamax$DepartType == "D"& Datamax$deparAge<= 1])/
                                           length( Datamax$deparAge[Datamax$DepartType == "D"]))%>%as.numeric(),
                                        median( Datamax$deparAge[Datamax$DepartType == "D"])%>%as.numeric(),
                                             mean( Datamax$deparAge[Datamax$DepartType == "D"])%>%as.numeric(),
                                             quantile( Datamax$deparAge[Datamax$DepartType == "D"],0.9)%>%as.numeric(),
                                             quantile( Datamax$deparAge[Datamax$DepartType == "D"],0.99)%>%as.numeric(),
                                             max( Datamax$deparAge[Datamax$DepartType == "D"])%>%as.numeric(),
                                             nrow(Datamax)
                                   )))
          }
        }
        metrics = rbind(metrics, 
                        tibble::tibble(Data = 'Raw_Longtresh',
                               firstage = MinAge,
                               param = c("S1m","S1y","L50", "MLE", "L90", "L99", "Max_longevity"),
                               stat = c("value"),
                               value = c(1-(length( data_sel$deparAge[data_sel$DepartType == "D"& data_sel$deparAge<= 1/12])/
                                           length( data_sel$deparAge[data_sel$DepartType == "D"]))%>%as.numeric(),
                                         1-(length(data_sel$deparAge[data_sel$DepartType == "D"& data_sel$deparAge<= 1])/
                                           length( data_sel$deparAge[data_sel$DepartType == "D"]))%>%as.numeric(),
                                        median(data_sel$deparAge[data_sel$DepartType == "D"])%>%as.numeric(),
                                         mean( data_sel$deparAge[data_sel$DepartType == "D"])%>%as.numeric(),
                                         quantile( data_sel$deparAge[data_sel$DepartType == "D"],0.9)%>%as.numeric(),
                                         quantile( data_sel$deparAge[data_sel$DepartType == "D"],0.99)%>%as.numeric(),
                                         max( data_sel$deparAge[data_sel$DepartType == "D"])%>%as.numeric()
                               )))
      }
      if ("Kaplan-Meier" %in% CalculateMetricsFrom){
        KMest = rawPLE%>%
          select(-Snd)%>%
          rename(Lx = ple)
        rawPLE$Z50 = (log(-log(rawPLE$ple))-log(-log(0.5)))*rawPLE$ple*log(rawPLE$ple)/(rawPLE$ple^2*rawPLE$Snd)
        rawPLE$Z10 = (log(-log(rawPLE$ple))-log(-log(0.1)))*rawPLE$ple*log(rawPLE$ple)/(rawPLE$ple^2*rawPLE$Snd)
        
        if(min(rawPLE$Z10, na.rm = T)>-1.96){
          mi10 = 1
        }else{
          mi10 = min(which(rawPLE$Z10<=-1.96))
        }
        if(max(rawPLE$Z10, na.rm = T)<1.96){
          ma10 = length(rawPLE$Z10)
        }else{
          ma10 = max(which(rawPLE$Z10>=1.96))
        }
        if(min(rawPLE$Z50, na.rm = T)>-1.96){
          mi50 = 1
        }else{
          mi50 = min(which(rawPLE$Z50<=-1.96))
        }
        if(max(rawPLE$Z50, na.rm = T)<1.96){
          ma50 = length(rawPLE$Z10)
        }else{
          ma50 = max(which(rawPLE$Z50>=1.96))
        }
        metrics = rbind(metrics,
                        tibble::tibble(Data = 'KM',
                               firstage = MinAge,
                               param = c(rep(c("L90", "L50"), each =3),"S1month", "S1year"),
                               stat = c(rep(c("mean", "lower", "upper"), 2),rep('value',2)),
                               value = c(
                                 KM_age(rawPLE, 0.1)[2]%>%as.numeric(), 
                                 rawPLE$Ages[ma10]%>%as.numeric(),
                                 rawPLE$Ages[mi10]%>%as.numeric(),
                                 KM_age(rawPLE, 0.5)[2]%>%as.numeric(),
                                 rawPLE$Ages[ma50]%>%as.numeric(),
                                 rawPLE$Ages[mi50]%>%as.numeric(),
                                 KM_Lx(rawPLE, 1/12)$Lx,KM_Lx(rawPLE, 1)$Lx%>%as.numeric()
                               )))
        rawPLE$Ndead = c(0,rawPLE$ple[1:(nrow(rawPLE)-1)] * 1000 -rawPLE$ple[2:nrow(rawPLE)]*1000)
        metrics = rbind(metrics,
                        tibble::tibble(Data = 'KM',
                               firstage = MinAge,
                               param = c("MLE", "H", "Epx", "G", "Ex", "CV"),
                               stat =rep("value",6),
                               value = c(
                                 sum(rawPLE$Ndead *rawPLE$Ages)/sum(rawPLE$Ndead)%>%as.numeric(), 
                                 CalcHx(rawPLE$ple,c(0,diff(rawPLE$Ages)) )%>%as.numeric(),
                                 -log(CalcHx(rawPLE$ple,c(0,diff(rawPLE$Ages))))%>%as.numeric(),
                                 CalcGx(rawPLE$ple,c(0,diff(rawPLE$Ages)) )%>%as.numeric(),
                                 CalcEx(rawPLE$ple,c(0,diff(rawPLE$Ages)))%>%as.numeric(),
                                 CalcCVx(rawPLE$ple,
                                         rawPLE$Ages,c(0,diff(rawPLE$Ages)))%>%as.numeric()
                               )))
      }
      
      # Run survival model using Basta --------------------------------------------
        # Extract BaSTA table:
      bastalist <- surv_Bastab(data_sel, DeathInformation = DeathInformation, EarliestDate = MinDate,
                               ExcludeStillBirth = TRUE)
      bastatab <- bastalist%>%
        dplyr::mutate(
          bdun = Max.Birth.Date-Min.Birth.Date,
          aliveTime = (Depart.Date - Entry.Date) / 365.25)%>%
        dplyr::filter(Depart.Date > Entry.Date)
      
      if(LastDead){
        bastatab$Depart.Type[bastatab$aliveTime == max(bastatab$aliveTime)]="D"
      }
      #Summary of data used
      summar$NBasta <- nrow(bastatab)
      summar$Ndead <- nrow(bastatab%>%dplyr::filter(Depart.Type =="D"))
      summar$Nrc <- nrow(bastatab%>%dplyr::filter(Depart.Type =="C"))
      summar$N8090 <- nrow(bastatab%>%dplyr::filter(year(Birth.Date) >= 1980 & year(Birth.Date) < 1990 ))
      summar$N9000 <- nrow(bastatab%>%dplyr::filter(year(Birth.Date) >= 1990 & year(Birth.Date) < 2000 ))
      summar$N0010 <- nrow(bastatab%>%dplyr::filter(year(Birth.Date) >= 2000 & year(Birth.Date) < 2010 ))
      summar$N1020 <- nrow(bastatab%>%dplyr::filter(year(Birth.Date) >= 2010 & year(Birth.Date) < 2020 ))
      summar$N2030 <- nrow(bastatab%>%dplyr::filter(year(Birth.Date) >= 2020))
      summar$QBD10 <- quantile(year(bastatab$Birth.Date),0.1)
      summar$QBD50 <- quantile(year(bastatab$Birth.Date),0.5)
      summar$QBD90 <- quantile(year(bastatab$Birth.Date),0.9)
      summar$BDincert<- as.numeric(mean(bastatab$Max.Birth.Date - bastatab$Min.Birth.Date))
      summar$Ndead <- nrow(bastatab%>%dplyr::filter(Depart.Type =="D"))
      summar$maxAge <- as.numeric(max(bastatab$Depart.Date - bastatab$Birth.Date, na.rm = TRUE))
      summar$maxAlive <- as.numeric(max(bastatab$Depart.Date - bastatab$Entry.Date, na.rm = TRUE))
      
      if ("Model" %in% CalculateMetricsFrom){
        if(summar$NBasta>0){
          #Check the percentage of individuals with known births
          Perbirthknown =  length(which(bastatab$bdun<32 & bastatab$Entry.Type=="B")) / 
            summar$NBasta
          if(Perbirthknown >= MinBirthKnown){
            #Check the number of Institutions
            Instb =  unique(data_sel$FirstHoldingInstitution[data_sel$AnimalAnonID %in% bastatab$AnimalAnonID])
            Instl =  unique(data_sel$LastHoldingInstitution[data_sel$AnimalAnonID %in% bastatab$AnimalAnonID])
            if(length(unique(c(Instb,Instl)))>=MinInstitution){
              #Check the number of individual selected
              if (summar$NBasta >= MinNSur) {
                if(summar$NBasta <= MaxNSur){
                  tempList <- list()
                  DICmods <- tibble::tibble(Models = Models,
                                    DIC = 0)
                  #Run the different basta models
                  for (imod in 1:length(Models)) {
                    print(Models[imod])
                    tempList[[Models[imod]]] <- BaSTA::basta(
                      bastatab, dataType = "census", shape = Shape, minAge = MinAge, 
                      model = Models[imod], parallel = TRUE, 
                      ncpus = ncpus, nsim = nchain,
                      niter = niter, burnin = burnin, thinning = thinning)
                    
                    if (!is.na( tempList[[Models[imod]]]$DIC[1])) {
                      DICmods$DIC[imod] <-  tempList[[Models[imod]]]$DIC["DIC"]
                    }
                  }
                  # If no model converged, run the models with more MCMC iterations
                  if (all(DICmods$DIC == 0)) {
                    print("more iterations")
                    for (imod in seq_along(Models)) {
                      print(Models[imod])
                      tempList[[Models[imod]]] <- BaSTA::basta(
                        bastatab, dataType = "census", shape = Shape, minAge = MinAge, 
                        model = Models[imod], parallel = TRUE,
                        ncpus = ncpus, nsim = nchain, 
                        niter = niter*4, burnin = burnin*4-3, thinning = thinning)
                      if (!is.na(tempList[[Models[imod]]]$DIC[1])) {
                        DICmods$DIC[imod] <-tempList[[Models[imod]]]$DIC["DIC"]
                      }
                    } 
                  }
                  
                  # Survival model outputs --------------------------------------------------
                  if (any(DICmods$DIC != 0)) {
                    a = which(DICmods$DIC == 0)
                    DICmods2 = DICmods
                    if(length(a)>0){
                      DICmods2 = DICmods2[-a,]
                    }
                    idModSel <- which(DICmods$DIC == min(DICmods2$DIC, na.rm = TRUE))
                    bastaRes <- tempList[[idModSel]]
                    summar$model = bastaRes$modelSpecs[["model"]]
                    summar$analyzed = TRUE
                  } else {
                    summar$error = 'no DIC from Basta' #nocov
                    summar$Nerr = 9 #nocov
                  }
                } else {
                  summar$error = "Nbasta > MaxNSur" #nocov
                  summar$Nerr = 8 #nocov
                }} else {
                  summar$error = "Nbasta < MinNSur" #nocov
                  summar$Nerr = 7 #nocov
                }
            }else{
              summar$error = "Number of institutions < MinInstitution" #nocov
              summar$Nerr = 6 #nocov
            }
          }else{ 
            summar$error = "%known births < MinBirthKnown" #nocov
            summar$Nerr = 5 #nocov
          }
        }else{
          summar$error = "NBasta = 0" #nocov
          summar$Nerr = 4 #nocov
        }}
    }else{
      summar$error = "lxMin > 0.75" #nocov
      summar$Nerr = 3 #nocov
    }
  } else {
    summar$error = "Nuncertdeath < MinNSur" #nocov
    summar$Nerr = 2 #nocov
  }
  
  return(list(summary = summar, metrics = metrics, bastaRes = bastaRes, 
              DICmods = DICmods, KM_estimator = KMest, bastatab = bastatab))
}

#Useful Functions
# life expectancy:
CalcEx <- function(Sx, dx) sum(Sx * dx) / Sx[1]

# Keyfitz's entropy:
CalcHx <- function(Sx, dx) {
  Sx1 <- Sx[Sx > 0]; Sx1 <- Sx1 / Sx1[1]
  if(length(dx)>1){dx <- dx[Sx > 0]}
  -sum(Sx1 * log(Sx1) * dx) / sum(Sx1 * dx)
}

# Gini coefficient:
CalcGx <- function(Sx, dx) {
  Sx <- Sx / Sx[1]
  if(length(dx)>1){dx <- dx[Sx > 0]}
  Sx <- Sx[Sx > 0]
  return(1 - 1 / sum(Sx * dx) * sum(Sx^2 * dx))
}

# Coefficient of variation:
CalcCVx <- function( Sx,x, dx) {
  Sx <- Sx / Sx[1]
  idd <- which(Sx > 0)
  Sx <- Sx[idd]
  dx <- dx[idd]
  x <- (x - x[1])[idd]
  dS <- -diff(Sx)
  dS <- dS / sum(dS)
  ex <- sum(Sx * dx)
  return(sqrt(sum((x[-length(x)] + dx[2:length(dx)]/2 - ex)^2 * dS)) / ex)
}

```

```{r examples-Sur_ana}
data(toy_core)
data(toy_deathinformation)
out <- Sur_ana(toy_core,  DeathInformation = toy_deathinformation, Models = "GO", Shape = "simple",
               niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
```

```{r tests-Sur_ana}
test_that("Sur_ana works", {
  data(toy_core)
  data(toy_deathinformation)
  out <- Sur_ana(toy_core, DeathInformation = toy_deathinformation, Models = "GO", Shape = "bathtub",
                 niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3,
                 LastDead = TRUE)
  expect_named(out, c("summary", "metrics", "bastaRes", "DICmods", 'KM_estimator', 'bastatab'))
  expect_named(out$DICmods, c("Models",  "DIC"))
  expect_equal(nrow(out$DICmods),1)
  expect_named(out$summary, c('NSelect', 'NUncertdeath', 'NBasta', 'Ndead', 'Nrc', 'N8090', 'N9000', 'N0010', 'N1020', 'N2030', 'QBD10', 'QBD50', 'QBD90', 'BDincert', 'maxAge', 'maxAlive', 'lxMin', 'outLev', 'model', 'analyzed', 'Nerr', 'error'))
  #expect_true(out$summary$analyzed )
  expect_true(is.numeric(out$summary$NSelect ))
  expect_true(is.numeric(out$summary$NBasta ))
  expect_true(is.numeric(out$summary$Ndead ))
  expect_true(is.numeric(out$summary$maxAge ))
  expect_true(is.numeric(out$summary$maxAlive))
  expect_true(is.numeric(out$summary$lxMin ))
  expect_true(is.numeric(out$summary$outLev ))
  expect_true(is.numeric(out$summary$Nerr ))
})


test_that("Sur_ana works2", {
  data(toy_core)
  data(toy_deathinformation)
  out <- Sur_ana(toy_core, DeathInformation = toy_deathinformation, Models = "GO", Shape = "bathtub",
                 niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3,
                 OutlLev1 = 100,MaxNSur=1000,
                 LastDead = TRUE)
  expect_named(out, c("summary", "metrics", "bastaRes", "DICmods", 'KM_estimator', 'bastatab'))
})
```




# Key survival metrics

`Sur_out()` estimates key metrics from the selected survival model and returns plots for the survival model. It includes the main following arguments:

* `out`: the output of the survival analysis (returns from `Sur_ana()`)
* `PlotDir`: Directory to save the plots of the best model

In addition to the outputs of the survival analysis, it returns:
* Key survival metrics estimated from either raw data, the Kaplan-Meier estimator or the selected survival model. Estimates are from birth or from age at sexual maturity.
* A check list
* the remaining life expectancy per age (relex_from0)
* the probability to live 5 years more (Sur5)
* Age-specific survival  (Sur1)
* The monthly survival  (Sur1m)


```{r function-Sur_out}
#' Create output and plots for the survival analysis
#' 
#'  estimates key metrics from the selected survival model and returns illustrating plots of the survival model.
#'
#' @param out \code{list} output of the survival analysis,  returns from `Sur_ana()`
#' @param MinAge \code{numeric} Age at which the survival analysis should start, in years.  see ?basta for more information.
#' @param MinMLE \code{numeric} Value used for checks. Minimum survivorship allowed at mean life expectancy. Between 0 and 1.
#' @param MaxLE \code{numeric} Value used for checks. Maximum remaining life expectancy at last observed age. In years.
#' @param MinLx  \code{numeric} Value used for longevity threshold and for checks. between 0 and 1. Minimum reached survivorship from the raw Kaplan-Meier analysis. This number avoids running survival analysis if there are too few dead individuals in the data. Lower is better.
#' @param MaxAge \code{numeric} Maximum possible age in years. Only used for model predictions. This argument is not used to select data.
#' @param ncpus  \code{numeric} Number of CPU cores to use. 
#' @param PlotDir \code{character} Directory to save the plots. If \code{NULL}, no plot is saved
#' @param PlotName \code{character} Name used to save the plots.
#' 
#' @return The output list of the survival analysis including, in addition:
#' * Key survival metrics including Mean life expectancy (MLE & Ex), median life
#' expectancy (L50) and  age at which 90% of the individual died (Longevity = L90)
#' estimated from raw data, the Kaplan-Meier estimator, and the survival model. Estimates
#' include ages from birth or from age at sexual maturity (if given). First year
#' survival, First month survival, Entropy (H and Epx = -log(H)), coefficient of
#' variation (CV) and Gini coefficient (G) are also estimated from the Kaplan-Meier
#' estimator and the survival model.
#' * Checks of the fit for the selected survival model:
#'     * Gof_KM_coeff1: Percentage of age points where the KM estimator is outside of the
#'     95% CI of the survivorship estimated from the survival model.
#'     * Gof_KM_coeff2: Maximum sum of same sign residuals between KM estimator and
#'     survival model
#'     * Gof_Martingal_KM: Value of the test that must be compared to a kolmogorov Smirnov table and test telling if the model fits
#'     * LxatMLE: Estimated survivorship at mean life expectancy
#'     * LEmaxOage: Life expectancy at max observed age
#'     * KMMinLx: Minimum survivorship reached by the Kaplan-Meier estimator
#' * The remaining life expectancy per age (relex_from0)
#' * The probability to live 5 years more (Sur5)
#' * Age-specific survival  (Sur1)
#' * Monthly survival  (Sur1m)
#' * Plots if requested
#'     * One showing the convergence of chains, together with estimated survivorship and
#'     survival and mortality rates.
#'     * The second showing remaining life-expectancy, and the age-specific probability
#'     to live 1 year more.
#' @export
#'
#' @importFrom grDevices pdf dev.off
#' @importFrom graphics lines
#'
#' @examples
Sur_out <- function(out,
                    MinAge = 0, MinMLE = 0.1, MaxLE = 2, MinLx = 0.1,
                    MaxAge = 120,
                    ncpus = 2, 
                    PlotDir = NULL, PlotName = ''
) {
  # Check correct format for inputs ---------------------------------------------
  assert_that(is.list(out))
  assert_that(out %has_name% c("summary", "metrics", "bastaRes", "DICmods", "bastatab"))
  
  assert_that(!is.null(out$summary))
  assert_that(!is.null(out$metrics))
  assert_that(!is.null(out$bastaRes),
              msg = "out$bastaRes must not be NULL. Sur_out() requires a fitted BaSTA model.")
  assert_that(!is.null(out$bastatab))
  
  assert_that(is.numeric(MinAge))
  assert_that(length(MinAge) == 1)
  assert_that(MinAge >= 0)
  
  assert_that(is.numeric(MaxAge))
  assert_that(length(MaxAge) == 1)
  assert_that(MinAge < MaxAge)
  
  assert_that(is.numeric(MinLx))
  assert_that(length(MinLx) == 1)
  assert_that(MinLx >= 0 && MinLx <= 1)
  
  assert_that(is.numeric(MinMLE))
  assert_that(length(MinMLE) == 1)
  assert_that(MinMLE >= 0 && MinMLE <= 1)
  
  assert_that(is.numeric(MaxLE))
  assert_that(length(MaxLE) == 1)
  
  assert_that(is.numeric(ncpus))
  assert_that(length(ncpus) == 1)
  assert_that(ncpus > 0)
  
  assert_that(is.character(PlotName))
  assert_that(length(PlotName) == 1)
  
  if (!is.null(PlotDir)) {
    checkmate::assert_directory_exists(PlotDir)
  }
  
  # Initialize outputs ----------------------------------------------------------
  out$check <- list(
    Gof_KM_coeff1 = NA_real_,
    Gof_KM_coeff2 = NA_real_,
    LxatMLE = NA_real_,
    LEmaxOage = NA_real_,
    KMMinLx = NA_real_,
    Gof_Martingal_KM = NA_real_,
    Gof_Martingal_test = NA
  )
  
  #Check if model is gompertz report b -----------------------------------------
  if(out$bastaRes$modelSpecs[["model"]] =="GO"){
    out$metrics = rbind(out$metrics,
                        tibble::tibble(Data = 'Model',
                               firstage = MinAge,
                               param = "b_Gomp",
                               stat = c("mean", "sd", "lower", "upper"),
                               value = out$bastaRes$coefficient[2,c(1:4)]
                        ))
  }
  
  # Define maximum predicted age: age at L(x) = 0.005---------------------------
  Lx_mean <- out$bastaRes$surv$nocov[1,]
  idMaxAge <- out$bastaRes$x[which(Lx_mean < 0.005)][1]+MinAge
  if (is.na(idMaxAge)) {
    MaxAge <- min(max(out$bastaRes$x)+MinAge, MaxAge)
  } else {MaxAge = idMaxAge}
  paramax = length(out$bastaRes$names)
  if("lambda" %in% out$bastaRes$names){paramax = paramax-1}
  
  
  # Estimate posterior of Lx from the best survival model-----------------------
  # run parallel estimation:
  dx = 0.01
  xv = seq(MinAge, MaxAge, by = dx)
    sfInit(parallel = TRUE, cpus = ncpus)
  on.exit(sfStop(), add = TRUE)
  Lxparal <- sfClusterApplyLB(1:ncpus, Sur_Lx, theMat =  out$bastaRes$params[,1:paramax],
                              model =  out$bastaRes$modelSpecs[["model"]], Shape =out$bastaRes$modelSpecs[["shape"]],  
                              iseq = floor(seq(0, nrow(out$bastaRes$params), length = ncpus + 1)),
                              xv = xv-MinAge)
  sfStop()
  # Gather estimates:
  for (jj in 1:ncpus) {
    if (jj == 1) {
      Lx <- Lxparal[[jj]]
    } else {
      Lx <- rbind( Lx, Lxparal[[jj]])
    }
  }
  
  # Estilmate key survival metrics from best model--------------------------------
  CVx = apply(Lx,1, CalcCVx, x = xv, dx = rep(dx,length(xv)))
  Hx = apply(Lx,1,CalcHx, dx =dx)
  out$metrics = rbind(out$metrics,
                      tibble::tibble(Data = 'Model',
                             firstage = MinAge,
                             param = rep(c("MLE", "H", "Epx", "G", "CV", "Hx"),each = 4),
                             stat = rep(c("mean", "sd", "lower", "upper"),6),
                             value = c(
                               out$bastaRes$PS$nocov$PS[1,]%>%as.numeric(),
                               out$bastaRes$PS$nocov$PS[2,]%>%as.numeric(),
                               out$bastaRes$PS$nocov$PS[3,]%>%as.numeric(),
                               out$bastaRes$PS$nocov$PS[4,]%>%as.numeric(), 
                               mean(CVx), sd(CVx), quantile(CVx, 0.025), quantile(CVx, 0.975) ,
                               mean(Hx), sd(Hx), quantile(Hx, 0.025), quantile(Hx, 0.975) 
                             )))
  # Remaining life expectancy:
  out$relex_from0<- Sur_relex(Lx, xv = xv, dx = dx)
  # Proba to live 5 year more:
  if(max(xv) >xv[1]+5){
    out$Sur5 <- Sur_age(Lx, xv = xv, Nyear = 5)
  }else{out$Sur5 = "Error : Age max lower than 5 years"} #nocov
  # Proba to live 1 year more:
  out$Sur1 <- Sur_age(Lx, xv = xv, Nyear = 1)
  # Proba to live 1 month more:
  out$Sur1m <- Sur_age(Lx, xv = xv, Nyear = 1/12)
  
  out$metrics = rbind(out$metrics,
                      tibble::tibble(Data = 'Model',
                             firstage = MinAge,
                             param = rep(c("remex0", "L50", "L90", "S1month", "S1year"),each =3),
                             stat = rep(c("mean", "lower", "upper"), 5),
                             value = c(
                               out$relex_from0[1, 2:4]%>%as.numeric(),
                               Sur_xx(Lx, xx = 0.5, xv = xv)%>%as.numeric(),
                               Sur_xx(Lx, xx = 0.1, xv = xv)%>%as.numeric(),
                               out$Sur1m[1, 2:4]%>%as.numeric(),
                               out$Sur1[1, 2:4]%>%as.numeric()
                             )))
  
  # Checks for selected model --------------------------------------------------
  #Check if the life expectancy at max observed age is below MaxLE years old
  MaxAgeobs = max(out$bastaRes$lifeTable$noCov$Mean$Ages)
  idMaxAge = ifelse(max(out$relex_from0$Age) > MaxAgeobs, which(out$relex_from0$Age==MaxAgeobs), length(out$relex_from0$Age))
  out$check$LEmaxOage = out$relex_from0$RemLExp[idMaxAge]

  
  #Check  if the survivorship at mean life expectancy is higher than MinMLE
  lx <-out$bastaRes$lifeTable$noCov$Mean$lx
  dif <-abs(out$bastaRes$lifeTable$noCov$Mean$Ages - out$bastaRes$PS$nocov$PS[1,1])
  out$check$LxatMLE = lx[which(dif == min(dif))]
  out$check$KMMinLx = out$summary$lxMin

  
  
  #Check if kaplan-meier estimator is within the 95% CI of the estimated Lx from the selected model
  roundN = ifelse(max(out$bastaRes$x+MinAge)>5, 2,3)
  Lx = tibble::tibble(Meann = out$bastaRes$surv$nocov[1,],
              Lower = out$bastaRes$surv$nocov[2,],
              Upper = out$bastaRes$surv$nocov[3,],
              Ages = round(out$bastaRes$x+round(MinAge,1),roundN))
  #correct estimates when MinAge >0
  Lx$Meann=Lx$Meann/Lx$Meann[1]
  Lx$Lower=Lx$Lower/Lx$Lower[1]
  Lx$Upper=Lx$Upper/Lx$Upper[1]
  
  LT = out$bastaRes$lifeTable$noCov$ple%>%
    dplyr::mutate(Ages = round(Ages/(5*10^(-roundN)))*(5*10^(-roundN)))%>%
    group_by(Ages)%>%
    summarise(ple = max(ple))%>%
    left_join(Lx)%>%
    tidyr::drop_na(Lower, Upper)%>%
    dplyr::mutate(check = (ple <= Upper & ple >= Lower),
           residual = Meann - ple,
           sign = ifelse(residual <0, -1,1),
           diff = 0)
  LT$diff[2:nrow(LT)] =ifelse(LT$sign[1:(nrow(LT)-1)] == LT$sign[2:nrow(LT)], 0,1)
  LT$diff = cumsum(LT$diff)
  v <- LT%>%
    group_by(diff)%>%
    summarise(resi = abs(sum (residual)))
  check = length(which(LT$check))/nrow(LT)
  out$check$Gof_KM_coeff1 =check
  out$check$Gof_KM_coeff2 =max(v$resi)
 
    #Martingale residuals ---------------------------------------------------------
  #REf = PRice and Jones 2017 bioRxiv A general goodness-of-fit test for survival analysis
  # Censored or dead
  delta = ifelse(out$bastatab$Depart.Type == "D", 1, 0)
  #censoring rate
  rho = length(which(out$bastatab$Depart.Type == "C"))/nrow(out$bastatab)
  # Add age at death and corresponding age id lines for join with Lx
  temp= out$bastatab%>%
    dplyr::mutate(agedeath = (Depart.Date - Birth.Date) / 365.25,
           Ages = round(agedeath/(5*10^(-roundN)))*(5*10^(-roundN)),
           Ages = round(Ages,roundN))
  if (max(temp$Ages)>max(Lx$Ages)) {
    temp = temp%>%
      dplyr::filter(Ages <= max(Lx$Ages))
  }
  temp = temp%>% rowwise()%>%
    dplyr::mutate(idage = which(Lx$Ages==Ages))
  # Martingale residuals
  D = 0.5*(1-(1+delta)* Lx$Meann[temp$idage])
  D = round(D,roundN)
  # Possible value for D
  d = seq(-0.5,0.5,0.01)
  #Observed distribution
  Gobs = cumsum(table(factor(D, levels = d))%>%as.numeric)
  Gobs = Gobs/max(Gobs)
  
  # theoretical probability distribution for martingale residuals
  gi = c((1/2-d[1:50])^(rho/(1-rho)),(1/2-d[51:101])^(rho/(1-rho))+2*rho/(1-rho)*(1-2*d[51:101])^(rho/(1-rho)))
  # theoretical Cumulative distribution
  G=cumsum(gi)/sum(gi)
  out$check$Gof_Martingal_KM = max(abs(Gobs-G))
  out$check$Gof_Martingal_test = out$check$Gof_Martingal_KM<= max(0.13, 1.358/sqrt(nrow(temp)))
  if(!out$check$Gof_Martingal_test){
    out$summary$error = "Gof martingale does not fit" #nocov
    out$summary$Nerr = 10 #nocov
  }
  
  # Plots------------------------------------------------------------------------
  if(!is.null(PlotDir)){
    pdf(file = paste0(PlotDir,"/",out$summary$Nerr,'_', PlotName, "_surcheck.pdf", sep=""),
        width = 6, height = 6)
    plot(out$bastaRes, main = PlotName)
    plot(out$bastaRes, plot.type = 'demorates')
    plot(out$bastaRes, plot.type = 'gof')
    dev.off()
    
    pdf(file = paste0(PlotDir,"/", out$summary$Nerr,'_', PlotName, "_surplot.pdf", sep=""), 
        width = 6, height = 6)
    plot(out$relex_from0$RemLExp~  out$relex_from0$Age, main = PlotName, xlab = "Age (year)", 
         ylab = "Remaining life expectancy", type = "l")
    lines(out$relex_from0$Lower ~  out$relex_from0$Age, lty = 2)
    lines(out$relex_from0$Upper ~  out$relex_from0$Age, lty = 2)
    
    plot(out$Sur1$Sur ~  out$Sur1$Age, 
         xlab = "Age (year)", ylab = 'Age-specific survival', type = "l")
    lines(out$Sur1$Lower ~  out$Sur1$Age, lty = 2)
    lines(out$Sur1$Upper ~  out$Sur1$Age, lty = 2)
    dev.off()
  }
  return(out)
}

```

```{r examples-Sur_out}
data(toy_core)
data(toy_deathinformation)
out <- Sur_ana(toy_core,  DeathInformation = toy_deathinformation, 
               Models = "GO", Shape = "simple",
               niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)

out <- Sur_out(out, ncpus = 3)
```

```{r tests-Sur_out}
test_that("Sur_out works", {
  PlotDir = paste0(tempdir(check = TRUE),'/temp')
  dir.create(PlotDir)
  data(toy_core)
  data(toy_deathinformation)
  out <- Sur_ana(toy_core,  DeathInformation = toy_deathinformation, Models = "GO", Shape = "simple",
                 niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
  out <- Sur_out(out, ncpus = 3, PlotDir = PlotDir)
  expect_named(out, c("summary","metrics",  "bastaRes", "DICmods", 'KM_estimator', "bastatab", "check", 'relex_from0', 'Sur5', 'Sur1', 'Sur1m'))
  expect_named(out$check, c('Gof_KM_coeff1', 'Gof_KM_coeff2', 'LxatMLE', 'LEmaxOage', 'KMMinLx', 'Gof_Martingal_KM' , 'Gof_Martingal_test'))
  expect_named(out$relex_from0, c("Age", "RemLExp", "Lower", "Upper"))
  expect_true(is.numeric(out$relex_from0$Upper))
  expect_true(is.numeric(out$relex_from0$Age))
  expect_true(is.numeric(out$relex_from0$RemLExp))
  expect_true(is.numeric(out$relex_from0$Lower))
  expect_true(out$relex_from0$Lower[1]<out$relex_from0$RemLExp[1])
  expect_true(out$relex_from0$RemLExp[1]<out$relex_from0$Upper[1])
  expect_named(out$Sur5, c("Age", "Lower",  "Upper", "Sur"))
  expect_true(is.numeric(out$Sur5$Lower))
  expect_true(is.numeric(out$Sur5$Sur))
  out <- Sur_out(out, ncpus = 3, PlotDir = PlotDir)
  
  unlink(PlotDir, recursive = TRUE)
})
```


# Create Basta data frame

This function formats the data prior to running BaSTA models. It checks column names and the succession of dates for each individual. It includes the following arguments:

* `coresubset` data frame including at least the following columns *AnimalAnonID*, *binSpecies*, *Class*, *Order*, *Family*, *CommonName*, *BirthDate*, *DepartDate*, *EntryDate*, *MaxBirthDate*, *MinBirthDate*, *EntryType*, and *DepartType*
* `DeathInformation` data.frame including at least the following columns *AnimalAnonID* and *RelevantDeathInformationType*
* `EarliestDate` Earlier date to be included.
* `LatestDate` Latest date to be included.
* `OtherCovars` Additional variables to include in the data
* `ExcludeStillBirth` Whether to exclude still births.

This function returns the subset dataset excluding individuals:
* with NA in the columns BirthDate, MinBirthDate, MaxBirthDate, EntryDate, and DepartDate.
* for which the dates from min birth date/entry date to Depart date do not follow one another in a logical order.
* Still born if required
* with depart date earlier than EarliestDate
* with entry date later than LatestDate
* Depart dates later than LatestDate are changed to latest date. These individuals are considered right-censored.



```{r function-surv_Bastab}
#'
#' Checks required columns and the logical succession of dates, then formats the
#' data for use in BaSTA.
#'
#' @param coresubset \code{data.frame} including at least the following columns:
#'   *AnimalAnonID*, *binSpecies*, *Class*, *Order*, *Family*, *CommonName*,
#'   *BirthDate*, *DepartDate*, *EntryDate*, *MaxBirthDate*, *MinBirthDate*,
#'   *EntryType*, and *DepartType*.
#' @param DeathInformation \code{data.frame} including at least the following
#'   columns *AnimalAnonID* and *RelevantDeathInformationType*.
#' @param EarliestDate \code{character} Earliest date to include, in
#'   `"YYYY-MM-DD"` format, or \code{NA}.
#' @param LatestDate \code{character} Latest date to include, in `"YYYY-MM-DD"`
#'   format, or \code{NA}.
#' @param OtherCovars \code{character} vector of additional variables to include.
#' @param ExcludeStillBirth \code{logical} Whether stillborn and fetal death
#'   records should be excluded.
#' 
#' @details
#' The function removes:
#' \itemize{
#'   \item individuals with missing values in `BirthDate`, `MinBirthDate`,
#'   `MaxBirthDate`, `EntryDate`, or `DepartDate`;
#'   \item individuals for which the sequence of dates is not logically consistent;
#'   \item stillborn or fetal death records, if requested;
#'   \item individuals with departure date earlier than `EarliestDate`;
#'   \item individuals with entry date later than `LatestDate`.
#' }
#'
#' Departure dates later than `LatestDate` are truncated to `LatestDate`, and the
#' corresponding individuals are treated as right-censored.
#' 
#' @return A \code{data.frame} formatted for BaSTA, with date columns renamed to
#'   `Birth.Date`, `Min.Birth.Date`, `Max.Birth.Date`, `Entry.Date`,
#'   `Depart.Date`, `Entry.Type`, and `Depart.Type`. Additional covariates are
#'   kept if requested.
#'   
#' @export
#' @importFrom tidyr replace_na
#'
#' @examples
surv_Bastab <- function (coresubset, DeathInformation, 
                         EarliestDate = NA, LatestDate = NA, 
                         OtherCovars = NA, ExcludeStillBirth = FALSE) 
{   
  
   # Check correct format for inputs --------------------------------------------
  inclcols <- c(
    "AnimalAnonID", "binSpecies", "Class", "Order", "Family", "CommonName",
    "BirthDate", "MinBirthDate", "MaxBirthDate", "EntryDate", "DepartDate",
    "EntryType", "DepartType"
  )
  
  assert_that(is.data.frame(coresubset))
  assert_that(coresubset %has_name% inclcols)
  
  assert_that(is.data.frame(DeathInformation))
  assert_that(DeathInformation %has_name% c("AnimalAnonID", "RelevantDeathInformationType"))
  
  assert_that(is.logical(ExcludeStillBirth))
  assert_that(length(ExcludeStillBirth) == 1)
  
  if (all(!is.na(OtherCovars))) {
    assert_that(is.character(OtherCovars))
    assert_that(length(OtherCovars) >= 1)
    assert_that(coresubset %has_name% OtherCovars)
    inclcols <- c(inclcols, OtherCovars)
  }
  
  if (!is.na(EarliestDate)) {
    EarliestDate <- lubridate::as_date(EarliestDate)
    assert_that(!is.na(EarliestDate),
                msg = "EarliestDate must be NA or a valid date in format 'YYYY-MM-DD'")
  }
  
  if (!is.na(LatestDate)) {
    LatestDate <- lubridate::as_date(LatestDate)
    assert_that(!is.na(LatestDate),
                msg = "LatestDate must be NA or a valid date in format 'YYYY-MM-DD'")
  }
  
  if (!is.na(EarliestDate) && !is.na(LatestDate)) {
    assert_that(EarliestDate <= LatestDate,
                msg = "EarliestDate must be earlier than or equal to LatestDate")
  }
  
  # Select columns and join RDI  -----------------------------------------------
  bastadat <- coresubset[, inclcols]%>%
    left_join(DeathInformation%>%select(AnimalAnonID, RelevantDeathInformationType)%>%
                dplyr::filter(RelevantDeathInformationType %in% c("Stillborn","Fetal death")), by = "AnimalAnonID")
  if (all(!is.na(OtherCovars))) {
    colnames(bastadat) <- c("AnimalAnonID", "binSpecies", "Class", 
                            "Order", "Family", "CommonName", "Birth.Date", "Min.Birth.Date", 
                            "Max.Birth.Date", "Entry.Date", "Depart.Date", "Entry.Type", 
                            "Depart.Type",OtherCovars, "RelevantDeathInformationType")
  }else{
    colnames(bastadat) <- c("AnimalAnonID", "binSpecies", "Class", 
                            "Order", "Family", "CommonName", "Birth.Date", "Min.Birth.Date", 
                            "Max.Birth.Date", "Entry.Date", "Depart.Date", "Entry.Type", 
                            "Depart.Type","RelevantDeathInformationType")
    
  }
  
  # Remove individuals with too Early or too late dates, and check logical succession of dates ------------------------
  if (is.na(EarliestDate)) {
    EarliestDate <- min(bastadat$Min.Birth.Date, na.rm = TRUE)
  }else{
    EarliestDate <- lubridate::as_date(EarliestDate)
  }
  if (is.na(LatestDate)) {
    LatestDate <- lubridate::today()
  }else{
    LatestDate <- lubridate::as_date(LatestDate)
  }
  
  bastadat <- bastadat%>%
    dplyr::mutate(Depart.Date = lubridate::as_date(Depart.Date),
           Entry.Date = lubridate::as_date(Entry.Date),
           Min.Birth.Date = lubridate::as_date(Min.Birth.Date),
           Max.Birth.Date = lubridate::as_date(Max.Birth.Date),
           Birth.Date = lubridate::as_date(Birth.Date)
    )%>%
    dplyr::filter(Depart.Date >= EarliestDate,
           Entry.Date <= LatestDate)%>%
    tidyr::drop_na(Birth.Date, Min.Birth.Date, Max.Birth.Date, Entry.Date, Depart.Date)%>%
    dplyr::filter((Min.Birth.Date <= Birth.Date)%>% replace_na(TRUE), 
           (Birth.Date <= Max.Birth.Date)%>% replace_na(TRUE), 
           (Birth.Date <= Entry.Date)%>% replace_na(TRUE), 
           (Entry.Date <= Depart.Date)%>% replace_na(TRUE))%>%
    dplyr::mutate(Depart.Type =dplyr:: if_else (Depart.Date > LatestDate, "C", Depart.Type),
           Depart.Date =dplyr:: if_else (Depart.Date > LatestDate, LatestDate, Depart.Date),
           Entry.Type  =dplyr:: if_else (Entry.Date < EarliestDate, "C", Entry.Type),
           Entry.Date  =dplyr:: if_else (Entry.Date < EarliestDate, EarliestDate, Entry.Date)
    )
  
  # Remove Stillborn if required ------------------------------------------------
  if(ExcludeStillBirth){
    bastadat <- bastadat%>%
      dplyr::filter((Depart.Date != Birth.Date)%>% replace_na(TRUE),
             stringr::str_detect(RelevantDeathInformationType, "Stillborn", negate = T)%>% replace_na(TRUE),
             stringr::str_detect(RelevantDeathInformationType, "Fetal death", negate = T)%>% replace_na(TRUE))
  } 
  
  bastadat <- bastadat%>%
    select(-"RelevantDeathInformationType")
  
  return(bastadat)
}
```

```{r examples-surv_Bastab}
data(toy_core)
data(toy_deathinformation)
out<- surv_Bastab(toy_core, DeathInformation = toy_deathinformation,
                  EarliestDate = '1990-01-01', LatestDate = '2020-12-31', 
                  OtherCovars = "SexType", ExcludeStillBirth = TRUE)


```

```{r tests-surv_Bastab}
test_that("surv_Bastab works", {
  data(toy_core)
  data(toy_deathinformation)
  out<- surv_Bastab(toy_core, DeathInformation = toy_deathinformation, LatestDate = '2020-12-31', 
                    OtherCovars = c('FirstHoldingInstitution','LastHoldingInstitution'), ExcludeStillBirth = TRUE)
  expect_named(out, c("AnimalAnonID", "binSpecies", "Class", "Order", "Family", "CommonName", "Birth.Date", "Min.Birth.Date", "Max.Birth.Date", "Entry.Date", "Depart.Date", 'Entry.Type', 'Depart.Type', 'FirstHoldingInstitution', 'LastHoldingInstitution'))
  expect_true(max(out$Depart.Date)<=lubridate::as_date('2020-12-31'))
})
```




# Estimate raw survivorship from life table

This function estimated the Kaplan-Meier table

```{r function-Sur_ple}
#' Kaplan-Meier estimator
#' 
#' Compute the Kaplan-Meier product-limit estimator from individual entry and departure ages.
#' 
#' @param Data \code{data.frame} including at least the following columns *deparAge*, *entryAge* (\code{date}), and *DepartType* 
#'
#' @return A data frame with columns:
#' * Ages: event times
#' * ple: Kaplan-Meier survivorship estimate
#' * event : number of deaths at each age
#' * Snd: Greenwood-type cumulative variance term used downstream
#' 
#' @export
#'
#' @examples
#' 
Sur_ple <- function(Data) {
  # Check correct format for inputs --------------------------------------------
  assert_that(is.data.frame(Data))
  assert_that(Data %has_name% c("deparAge", "DepartType", "entryAge"))
  
  # Kaplan-Meier Estimator -----------------------------------------------------
  # Find records with same first and last age:
  Data <- Data%>%
    rowwise()%>%
    dplyr::mutate(idsame = entryAge == deparAge,
           deparAge = ifelse(deparAge < 0.1,0.1, deparAge),
           deparAge = round(ifelse(idsame, deparAge + 1/365.25, deparAge),1))%>%
    arrange(deparAge)
  
  # Number of ages:
  nage <-unique(Data$deparAge)
  
  # Cx and delta:
  Cx <- delx <- Snd <- rep(0, length(nage))
  
  # Fill up Cx and delta:
  ii = 0
  for (agev in nage) {
    ii = ii + 1
    idNx <-Data %>%dplyr::filter(entryAge <= agev,
                          deparAge >= agev)
    delx[ii] <- length(which(Data$DepartType == "D"  & Data$deparAge ==agev))
    Cx[ii] <- 1- delx[ii]/nrow(idNx)
    if(ii==1){
      Snd[ii]=delx[ii]/(nrow(idNx)*(nrow(idNx)-+delx[ii]))
    }else{
      Snd[ii]=Snd[ii-1]+delx[ii]/(nrow(idNx)*(nrow(idNx)-+delx[ii]))
      if((nrow(idNx)*(nrow(idNx)-+delx[ii])) == 0)  Snd[ii]=Snd[ii-1]
    }
  }
  
  # Calculate product limit estimator:
  ple <- cumprod(Cx)
  ple[ple<0]=0
  
  # Add age 0:
  if (nage[1] > 0) {
    nage <- c(0, nage)
    ple <- c(1, ple)
    delx<- c(0, delx)
    Snd<- c(0, Snd)
  }
  
  # Output:
  pleTab <- data.frame(Ages = nage, 
                       ple = ple, event = delx,  
                       Snd = Snd )
  
  return(pleTab)
}

```

```{r examples-Sur_ple}
entryAge = sample(c(1:10), 200, replace = TRUE)
data <- data.frame(
  entryAge = entryAge,
  deparAge =  entryAge + sample(c(0:10), 200, replace = TRUE),
  DepartType = sample(c('C', 'D'), 200, replace = TRUE))


out<-Sur_ple(data)
```

```{r tests-Sur_ple}
test_that("Sur_ple works", {
  entryAge = sample(c(1:10), 200, replace = T)
  data <- data.frame(
    entryAge = entryAge,
    deparAge =  entryAge + sample(c(0:10), 200, replace = T),
    DepartType = sample(c('C', 'D'), 200, replace = T))
  out<-Sur_ple(data)
  expect_named(out, c("Ages", "ple", "event", "Snd"))
  expect_true(min(out$ple)>=0)
  expect_true(max(out$ple)<=1)
})
```


# Remaining life expectancy

`Sur_relex()` estimates the remaining life expectancy over ages. It includes the following arguments: 

* `Lx` Survivorship
* `dx` precision for age (delta between successive age)
* `xv` Age vector

```{r function-Sur_relex}
#' Remaining life expectancy
#' 
#' Estimate remaining life expectancy over ages.
#' 
#'
#' @param Lx \code{numeric} Matrix of posterior survivorship trajectories
#' @param dx \code{numeric} Precision for age
#' @param xv \code{numeric} Age vector (same length as Lx)
#'
#' @return a data frame including age, the mean and 95% credible interval of the remaining life expectancy
#' 
#' @export
#' 
#' @importFrom paramDemo CalcSurv
#' @importFrom snowfall sfInit sfLibrary sfClusterApplyLB  sfStop
#' 
#' @examples
Sur_relex <- function(Lx, dx = 0.01,xv) {
  assert_that(is.numeric(dx))
  assert_that(dx > 0)
  assert_that(is.numeric(Lx))
  Sur_relex_0 <- function( Sx, dx = 0.01) {
    ex <- Sx * 0
    idn0 <- which(Sx > 0.001)
    ex[idn0] <- rev(cumsum(rev(Sx[idn0] * dx))) / Sx[idn0]
    return(ex)
  }
  
  exMat <-apply(Lx,1, Sur_relex_0,  dx = dx)
  exQuants <- data.frame(Age = xv, 
                         RemLExp = apply(exMat, 1,quantile, 0.5), 
                         Lower = apply(exMat, 1, quantile, 0.025),
                         Upper = apply(exMat, 1, quantile, 0.975))
  return(exQuants)
}

```

```{r examples-Sur_relex}
Lx = matrix(c(seq(1,0,by = -0.1), 1,seq(0.5,0,length.out = 10)),nrow =2)
out <- Sur_relex(Lx, dx = 1, xv = c(0:10))
```

```{r tests-Sur_relex}
Lx = matrix(c(seq(1,0,by = -0.1), 1,seq(0.5,0,length.out = 10)),nrow =2)
test_that("Sur_relex works", {
  out <- Sur_relex(Lx, dx = 1, xv = c(0:10))
  expect_named(out, c("Age", "RemLExp", "Lower", "Upper"))
  expect_true(is.numeric(out$RemLExp))
})
```



# Age-specific survival

`Sur_age()` estimates age-specific survival. It includes the following arguments: 

* `Lx` Survivorship
* `xv`  Age vector
* `Nyear` Number of year to survive. Default = 1

```{r function-Sur_age}
#' Age-specific survival
#' 
#' Estimate age-specific survival, i.e. the probability of surviving at least Nyear additional years, from the parameters outputs of a Basta model.
#' 
#'
#' @param Lx \code{numeric} Matrix of posterior survivorship trajectories
#' @param xv \code{numeric}  Age vector
#' @param Nyear \code{numeric} number of years to survive Default = 1#
#'
#' @return a data frame including age, the mean and 95% credible interval of the age specific probability to survive \code{Nyear} years.
#' 
#' @export
#' 
#' @importFrom paramDemo CalcSurv
#' @importFrom snowfall sfInit sfLibrary sfClusterApplyLB  sfStop
#' 
#' @examples
Sur_age <- function(Lx, xv, Nyear = 1) {
  assert_that(is.numeric(Nyear))
  assert_that(Nyear > 0)
  assert_that(is.numeric(xv))
  assert_that(all(xv >= 0))
  Sur_age_0 <- function(Sx, Nyear = 1, 
                        xv = seq(0, 50, by = 0.01)
  ) {
    
    dep = min(which(xv >= (xv[1]+Nyear)))
    idn0 <- length(Sx)
    ex = Sx[dep:idn0]/Sx[1:(idn0-dep+1)]
    return(ex)
  }
  
  exMat <- apply(Lx, 1, Sur_age_0, Nyear = Nyear,xv = xv )
  exMat[is.nan(exMat)]=0
  exQuants <- tibble::tibble(Age = xv[1:nrow(exMat)] , 
                     Lower = apply(exMat, 1, quantile, 0.025),
                     Upper = apply(exMat, 1, quantile, 0.975))%>%
    dplyr::mutate(Sur = apply(exMat, 1, mean))
  return(exQuants)
}

```

```{r examples-Sur_age}
Lx = matrix(c(seq(1,0,by = -0.1), 1,seq(0.5,0,length.out = 10)),nrow =2)
out <- Sur_age(Lx, Nyear = 5, xv = c(0:10))
```

```{r tests-Sur_age}
test_that("Sur_age works", {
  Lx = matrix(c(seq(1,0,by = -0.1), 1,seq(0.5,0,length.out = 10)),nrow =2, byrow =T)
  out <- Sur_age(Lx,  Nyear = 5, xv = c(0:10))
  expect_named(out, c("Age", "Lower", "Upper", "Sur"))
  expect_true(is.numeric(out$Sur))
  expect_equal(nrow(out), 6)
})
```

# Survivorship quantile
`Sur_xx()` estimates the age at which xx% of the population is still alive. It includes the following arguments: 

* `Lx` Survivorship
* `xv` Age vector
* `xx` Proportion of the population still alive

```{r function-Sur_xx}
#' Survivorship quantile
#' 
#' Estimate the age at which survivorship reaches a specified level xx
#' 
#' @param Lx \code{numeric} Survivorship
#' @param xv \code{numeric} Age vector
#' @param xx  \code{numeric} Proportion of the population still alive. Between 0 and 1.
#'
#' @return a list including the mean and 95% credible interval age at which xx% of the population is still alive
#' 
#' @export
#' 
#' @importFrom paramDemo CalcSurv
#' @importFrom snowfall sfInit sfLibrary sfClusterApplyLB  sfStop
#' 
#' @examples
Sur_xx <- function(Lx, xv, xx = 0.5) {
  assert_that(is.numeric(xx))
  assert_that(xx >= 0)
  assert_that(xx <= 1)
  
  Sur_xx_0 <- function(Sx, xx =0.5, xv = seq(0, 50, by = 0.01)){
    return( xv[max(which(Sx>=xx))])}
  
  exMat <- apply(Lx,1, Sur_xx_0, xv = xv , xx=xx)
  
  out <- list(L = mean(exMat), Lower = quantile(exMat,0.025), Upper =  quantile(exMat,0.975))
  
  return(out)
}

```

```{r examples-Sur_90}
Lx = matrix(c(seq(1,0,by = -0.1), 1,seq(0.5,0,length.out = 10)),nrow =2)
out <- Sur_xx(Lx,  xv = c(0:10), xx = 0.1)
```

```{r tests-Sur_90}
test_that("Sur_xx works", {
  Lx = matrix(c(seq(1,0,by = -0.1), 1,seq(0.5,0,length.out = 10)),nrow =2)
  out <- Sur_xx(Lx, xv = c(0:10), xx = 0.1)
  expect_named(out, c("L", "Lower", "Upper"))
  expect_true(is.numeric(out$L))
})
```

# Survivorship
`Sur_Lx()` estimates the survivorship from parameter of the survival model. It includes the following arguments: 

* `sim` iteration number if the function is run in parallel
* `theMat` The posteriors estimates of the model parameter
* `model` Name of the basta model: "GO", "EX", "LO" and/or "WE".
* `Shape` character Shape of the basta model: "simple", "Makeham", "bathtub". 
* `iseq` rows of themat to consider
* `xv`  Age vector


```{r function-Sur_Lx}
#' Estimate Lx
#' 
#' @param sim \code{numeric} iteration number if the function is run in parallel
#' @param theMat \code{array} The posteriors estimates of the model parameter
#' @param model \code{character}  Name of the basta model: "GO", "EX", "LO" and/or "WE".
#' @param Shape \code{character}  Shape of the basta model: "simple", "Makeham", or "bathtub". 
#' @param iseq \code{numeric}  index boundaries defining the rows processed by each parallel worker
#' @param xv \code{numeric}  Age vector
#'
#' @return Lx
#' 
#' @importFrom paramDemo CalcSurv
#' 
#'
#' @examples
#' @noRd
Sur_Lx <- function(sim= 1, theMat ,model = 'GO', Shape = 'bathtub',  
                   iseq = 1:nrow(theMat), 
                   xv = seq(0, 50, by = 0.01)
) {
  
  idseq <- (iseq[sim] + 1):iseq[sim + 1]
  surage <- t(sapply(idseq, function(ith) {
    theta <- theMat[ith, ]
    Sx <- paramDemo::CalcSurv(theta = theta, x = xv, model = model, shape = Shape)
    return(Sx)
  }))
  return(surage)
}
```


# Age from Kaplan-Meier

`KM_age()` estimates Ages at which Lx equals given values from the Kaplan-Meier curve. It includes the following arguments: 

* `KM_tab`: The kaplan-Meier table including the columns: `Ages` and `ple`.
* `Lx`: Values of Lx for which age is requested

```{r function-KM_age}
#'  Age from Kaplan-Meier
#' 
#' Estimate Ages at which Lx equals given values from the Kaplan-Meier table. Ages are linearly interpolation between adjacent KM.
#' 
#' @param KM_tab \code{data.frame} The kaplan-Meier table including the columns: *Ages* and *ple*
#' @param Lx \code{numeric vector} Values of Lx for which age is requested
#'
#' @return a data frame including Lx and Age
#' 
#' @export
#' 
#' @importFrom paramDemo CalcSurv
#' @importFrom snowfall sfInit sfLibrary sfClusterApplyLB  sfStop
#' 
#' @examples
KM_age <- function(KM_tab, Lx) {
  
  assert_that(is.data.frame(KM_tab))
  assert_that(KM_tab %has_name% c("Ages", "ple"))
  assert_that(is.numeric(Lx))
  assert_that(all(Lx > 0))
  assert_that(all(Lx < 1))
  
  Age_out = c()
  for (lx in Lx){
    if(min(KM_tab$ple)>lx){
      Age_out =  c(Age_out,NA)
    }else{
      a = max(which((KM_tab$ple-lx) >=0))
      if (KM_tab$ple[a]== lx){
        Age_out =  c(Age_out,KM_tab$Ages[a])
        
      }else{
        #Linear extrapolation
        slope = (KM_tab$ple[a]-KM_tab$ple[a+1])/(KM_tab$Ages[a]-KM_tab$Ages[a+1])
        int = KM_tab$ple[a] - slope *(KM_tab$Ages[a])
        Age_out =  c(Age_out,(lx -int)/ slope)
      }
    }
  }
  
  
  out <- tibble::tibble(Lx = Lx, Age = Age_out)
  
  return(out)
}

```

```{r examples-KM_age}
KM_tab = data.frame( Ages = 1:10,
                     ple = sort(runif(10, 0, 1), decreasing = TRUE))
out <- KM_age(KM_tab, Lx = c(0.5,0.1))
```

```{r tests-KM_age}
KM_tab = data.frame( Ages = 1:10,
                     ple = sort(runif(10, 0, 1), decreasing = TRUE))
KM_tab2 = data.frame( Ages = 1:10,
                      ple = seq(1,0.1, by = -0.1))

test_that("KM_out works", {
  out <- KM_age(KM_tab, Lx = c(0.5,0.1))
  expect_named(out, c("Lx", "Age"))
  expect_true(is.numeric(out$Lx))
  expect_true(is.numeric(out$Age))
  out <- KM_age(KM_tab2, Lx = c(0.5,0.1))
})
```





# Lx from Kaplan-Meier

`KM_Lx()` estimates Lx at given ages from the Kaplan-Meier table. It includes the following arguments: 

* `KM_tab`: The kaplan-Meier table including the columns: `Ages` and `ple`.
* `Age`: Age for which Lx is requested

```{r function-KM_Lx}
#'  Lx from Kaplan-Meier
#' 
#' Estimate Survivorship  at given ages from the Kaplan-Meier table.
#' 
#' @param KM_tab \code{data.frame} The kaplan-Meier table including the columns: *Ages* and *ple*
#' @param Age \code{vector of numeric} Ages for which Lx is requested
#'
#' @return a data frame including age, and Lx, giving the interpolated Kaplan-Meier survivorship at the requested ages.
#' @export
#' 
#' @importFrom paramDemo CalcSurv
#' @importFrom snowfall sfInit sfLibrary sfClusterApplyLB  sfStop
#' 
#' @examples
KM_Lx <- function(KM_tab, Age) {
  
  assert_that(is.data.frame(KM_tab))
  assert_that(KM_tab %has_name% c("Ages", "ple"))
  assert_that(is.numeric(Age))
  assert_that(all(Age > 0))
  
  Lx_out = c()
  for (age in Age){
    if(max(KM_tab$Ages)<age){
      Lx_out =  c(Lx_out,NA)
    }else{
      l = min(which((KM_tab$Ages-age) >=0))
      if (KM_tab$Ages[l]== age){
        Lx_out =  c(Lx_out,KM_tab$ple[l])
        
      }else{
        #Linear extrapolation
        slope = (KM_tab$ple[l]-KM_tab$ple[l-1])/(KM_tab$Ages[l]-KM_tab$Ages[l-1])
        int = KM_tab$ple[l] - slope *(KM_tab$Ages[l])
        Lx_out =  c(Lx_out,age*slope +int)
      }
    }
  }
  out <- tibble::tibble(Age = Age, Lx = Lx_out)
  
  return(out)
}

```

```{r examples-KM_Lx}
KM_tab = data.frame( Ages = 1:10,
                     ple = sort(runif(10, 0, 1), decreasing = TRUE))
out <- KM_Lx(KM_tab, Age = c(3,6.5))
```

```{r tests-KM_Lx}
KM_tab = data.frame( Ages = 1:10,
                     ple = sort(runif(10, 0, 1), decreasing = TRUE))

test_that("KM_Lx works", {
  out <- KM_Lx(KM_tab, Age = c(3,6.5,12))
  expect_named(out, c("Age", "Lx"))
  expect_true(is.numeric(out$Lx))
  expect_true(is.numeric(out$Age))
})
```



# Gap analysis in longevity

`select_Longthreshold()` runs a gap analysis on the distribution of longevity and gives the threshold value to use to avoid having gaps in longevity. It can focus on a given sex using the following arguments:

* `Data` including at least the following columns *Birth.Date*, *Depart.Date*, *Entry.Date*, and *SexType*
* `SexCats` Male, Female or All
* `MinN` Minimum number of individuals to run the gap analysis
* `PlotDir` Directory to save the plots
* `PlotName` name of the graph to be saved

It returns the data with the selected sex and additional columns showing which individuals are above the percentiles 95%, 99% and 99.9%
and a summary list including the threshold value selected for the distribution of longevity. It plots the distribution of longevity with gaps if `PlotDir` is given. 

```{r function-select_Longthreshold}
#' Gap analysis in longevity
#' 
#' Run a gap analysis on the distribution of longevity and give the threshold value to use to avoid having gaps in longevity.
#' 
#' @param Data \code{data.frame} including at least the following columns *BirthDate* (\code{date}), *DepartDate* (\code{date}), *EntryDate* (\code{date}), and *SexType*
#' @param SexCats \code{character} Male, Female or All.
#' @param MinN \code{numeric} Minimum number of individuals to run the gap analysis.
#' @param PlotDir \code{character} Directory to save the plots.
#' @param PlotName \code{character} name of the graph to be saved.
#'
#' @return A list including
#' * the data with the selected sex and additional columns showing which individuals are above the percentiles 95%, 99% and 99.9%
#' * A summary list with:
#'     * Sex = the sex selected
#'     * Nselect: the number of individuals of this sex
#'     * Nlifespan : the number of individuals with estimated lifespan (i.e. estimated birth dates, censored individuals are also included)
#'     * GapThresh : The threshold value selected for the distribution of longevity
#'     * NThres : the number of individuals selected using this threshold
#' * Plot and save the longevity distribution with gaps if PlotDir is given.
#'  
#' @export
#'
#' @examples
select_Longthreshold <- function(Data,  SexCats = "All", MinN = 50,
                                 PlotDir = NULL, PlotName = ''
                                 ) {
  # Check correct format for inputs -----------------------------------------------------------------------
  assert_that(is.data.frame(Data))
  assert_that(Data  %has_name% c("BirthDate", "DepartDate",
                                 "EntryDate", "SexType"))
  assert_that(is.character(SexCats))
  assert_that(length(SexCats)==1, 
              msg = "You can chose only one sex category")
  assert_that(all(SexCats %in% c("Female", "Male", "All")))
  assert_that(is.numeric(MinN))
  if(!is.null(PlotDir)){
    assert_that(is.character(PlotDir))
    checkmate::assert_directory_exists(PlotDir)}
  
    # Initialize Output list_______________________________________________________
  outTab <- tibble::tibble( 
    Sex = SexCats,
    Nselect = 0,
    Nlifespan = 0,
    GapThresh = NA, 
    NThres = 0
  ) 
  
  #Calculate longevity and time alive ------------------------------------------
  coresubset <- Data%>%
    dplyr::mutate(
      #longevities:
      lifespans = as.numeric(DepartDate - BirthDate) / 365.25,
      # Calculate time alive:
      alive = as.numeric(DepartDate - EntryDate) / 365.25)
  

  # Select Sex -----------------------------------------------------------------
  if(SexCats %in% c('Male', 'Female')){
    coresex <- coresubset%>%dplyr::filter(SexType ==SexCats)
  }else{coresex = coresubset}
  outTab$Nselect = nrow(coresex)
  
  # Run Gap analysis -----------------------------------------------------------
  if(nrow(coresex)>0){
     quant = quantile (coresex$lifespans,c(0.95,0.99,0.999))
    coresex <-  coresex%>%
      dplyr::mutate(`above95`=dplyr:: if_else(lifespans> quant[1], 1, 0),
             `above99`=dplyr:: if_else(lifespans> quant[2], 1, 0),
             `above99.9`=dplyr:: if_else(lifespans> quant[3], 1, 0)
      )
    # Find gaps in longevities:
    corelong <- coresex%>%tidyr::drop_na(lifespans)
    outTab$Nlifespan <- nrow(corelong)
     if (nrow(corelong) > MinN) {
      gapsAlive <- find_gaps(corelong$alive, 
                             MaxAlive = quantile(corelong$alive, 0.5, na.rm = TRUE), 
                             plot = F)
      if (nrow(gapsAlive) > 0) {
        MaxAlive <-  gapsAlive$iniAge[1]
      } else {
        MaxAlive <- max(corelong$alive, na.rm = TRUE)
      }
      if (MaxAlive < 5) {
        MaxAlive <- 5 # nocov
      } else if (MaxAlive > 50) {
        MaxAlive <- 50
      }
      
      # Plot the longevity gaps
      pdf(file = glue::glue("{PlotDir}/{PlotName}_LongThres.pdf"), width = 5, height = 5)
      par(mar = c(4, 4, 1, 1))
      gaps <- find_gaps(corelong$lifespans, MaxAlive = MaxAlive, plot = T,
                        main = PlotName, 
                        xlab = "")
      ngap = nrow( gaps)
      if (nrow(gaps) > 0) {
        allev <- 0
        while(allev < 3 & ngap>0) {
          allev <- allev + 1
          qlev <- c("99.9", "99", "95")[allev]
          abcol <- sprintf("above%s", qlev)
          gaps <- find_gaps(corelong$lifespans[corelong[[abcol]] == 0], 
                            MaxAlive = MaxAlive, plot = T,
                            main = paste(qlev, "%"), 
                            xlab = "")
          ngap = nrow(gaps)
        }
        outTab$GapThresh <-as.numeric(qlev)
          if (allev < 3) {
          for (ll in (allev + 1):3) {
            qlev <- c("99.9", "99", "95")[ll]
            abcol <- sprintf("above%s", qlev)
            plot(c(0, 1), c(0, 1), col = NA, axes = FALSE, xlab = "", 
                 ylab = "", main = paste(qlev, "%"))
            text(0.5, 0.5, "No gaps", cex = 2)
            box()
          }
        }
      } else {
        outTab$GapThresh <-100 # nocov
      }
      outTab$NThres <- nrow(coresex)
      dev.off()
    } 
  }
  return(list(summar = outTab, data = coresex) )
}
```

```{r examples-select_Longthreshold}
TempDir <- paste0(tempdir(check = TRUE),'/temp')
dir.create(TempDir)
data(toy_core) #### CHANGE DATASET WITH ONE EXCLUDING ABOVE95 99 99.9
out <- select_Longthreshold (Data = toy_core,  SexCats = "All", 
                             PlotDir = TempDir, PlotName = "Testudo_hermanni")
list.files(TempDir)
#reMove temporary folder
unlink(TempDir, recursive = TRUE)
```

```{r tests-select_Longthreshold}
test_that("select_Longthreshold works", {
  TempDir <- paste0(tempdir(check = TRUE),'/temp')
  dir.create(TempDir)
  data(toy_core) #### CHANGE DATASET WITH ONE EXCLUDING ABOVE95 99 99.9
  out <- select_Longthreshold (Data = toy_core,  SexCats = "All", 
                               PlotDir = TempDir, PlotName = "Testudo_hermanni")
  expect_true(file.exists(paste(TempDir, 'Testudo_hermanni_LongThres.pdf', sep = '/')))
  #reMove temporary folder
  expect_named(out, c("summar",  "data"))
  expect_true("above95" %in% names(out$data))
  expect_named(out$summar, c("Sex", 'Nselect', "Nlifespan", "GapThresh", "NThres"))
  expect_true(out$summar$Sex == "All" )
  expect_equal(out$summar$Nselect, nrow(toy_core))
  out <- select_Longthreshold (Data = toy_core[1:30,],  SexCats = "Female", 
                               PlotDir = TempDir, PlotName = "Testudo_hermanni")
  unlink(TempDir, recursive = TRUE)

})
```


## Find gaps in a continuous variable

`find_gaps` Looks for gaps within a continuous variable.Values of the variable are rounded and gaps are found if one integer is not represented. Gaps are returned only if the length of the gaps is 20% higher than the starting age of this gap. It uses the following arguments:

* `x` variable to analyze
* `MaxAlive` Return gaps higher than this maximum, only. Default = NA

It returns a data frame where each line is a gap described by its initial/starting age, its final/ending age and its length

```{r function-find_gaps}
#' Find gaps in an integer variable
#' 
#' Looks for gaps within a continuous variable. Values of the variable are rounded and gaps are found if one integer is not represented. Gaps are returned only if the length of the gaps is 20% higher than the starting age of this gap.
#'
#' @param x \code{vector of numeric} variable to analyze
#' @param MaxAlive \code{numeric} Return gaps higher than this maximum, only.
#' @param plot \code{logical} whether to actually plot the histogram
#' @param ... arguments pass to hist()
#'
#' @return
#' A data frame where each line is a gap described by its initial/starting age, its final/ending age and its length
#' 
#' @export
#'
#' @importFrom graphics abline box hist par text
#'
#' @examples
find_gaps <- function(x,MaxAlive = NA, plot = FALSE,...) {
   # Check correct format for inputs -----------------------------------------------------------------------
 assert_that(is.numeric(x))
  if(is.na(MaxAlive)){
    MaxAlive = max(x)
  }else{
    assert_that(is.numeric(MaxAlive))}
  assert_that(is.logical(plot))
  
  #Round each value of the continuous variable-----------------------------------
  breaks <- floor(min(x, na.rm = TRUE)):ceiling(max(x, na.rm = TRUE))
  hs <- hist(x, breaks = breaks, plot = plot, ...)
  xc <- hs$counts
  xc[hs$counts > 0] <- 1
  
  # Look for gaps ---------------------------------------------------------------
  fin <- FALSE
  i <- 1
  n <- length(xc)
  gaps <- tibble::tibble(iniAge= numeric(1), finAge= numeric(1), Gap= numeric(1))
  while(!fin & i < n) {
    id0st <- which(xc[i:n] == 0)[1]
    if (!is.na(id0st)) {
      id0fin <- which(xc[(i + id0st - 1):n] == 1)[1] - 1
      gaps <- rbind(gaps, c(hs$breaks[i + id0st - 1], 
                            hs$breaks[i + id0st + id0fin - 2], id0fin))
      i <- i + id0st + id0fin - 1
    } else {
      fin <- TRUE
    }
  }
 
  #Return gaps only if the gaps is 20% higher than the starting age of this gap-----
  idgap <- gaps%>%
    dplyr::filter(Gap/iniAge > 0.2 ,
           finAge  >= MaxAlive)
   if (nrow(idgap) > 0 & plot) {abline(v = idgap$iniAge, col = 2)}
  return(idgap)
}
```

```{r examples-find_gaps}
x = runif(10,0,40)
out<-find_gaps(x,MaxAlive = 5, plot = FALSE)
```

```{r tests-find_gaps}
test_that("find_gaps works", {
  x = runif(10,0,40)
  out<-find_gaps(x,MaxAlive = 5, plot = FALSE)
  out<-find_gaps(x,MaxAlive = NA, plot = FALSE)
  expect_named(out,c('iniAge', 'finAge', 'Gap'))
  expect_equal(out$finAge-out$iniAge+1, out$Gap)
})
```



```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_select.Rmd", vignette_name = "Data Selection")
```




```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_survival.Rmd", vignette_name = "Survival")
```

